---
title: "Final Figures and Statistics for Publication"
author: "Callie Hundley"
date: '2026-09-12'
output: html_document
---

```{r, include = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, size = 'scriptsize')
```

```{r wrap-hook, eval=FALSE, include=FALSE}
# Define a new code chunk option 'linewidth' for wrapping unruly R output
hook_output = knitr::knit_hooks$get('output')
knitr::knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n, exdent = 3, indent = 3)
    x = paste(x, collapse = '\n')} hook_output(x, options)})

```


# Load packages 

```{r}
#Load Packages
library(readr)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(survival)
library(coxme) 
library(car)
library(tidyr)
library(dplyr)
library(survminer)
library(writexl)

## Colleen added:
library(performance)
library(lme4)
library(lmerTest)
library(emmeans)

```

This **R markdown document** outlines the data sets and codes used to create all the figures and run statistics published in the paper. Data associated with all physiological metrics (everything besides mortality) recorded from T4 (the timeline associated with this paper) was compiled into the file `masterdata_colony.csv` alongside older time points for long-term storage. Therefore, some analysis sections will be calling upon this file/dataset. But for the physiological metrics that were assessed using lab-based analyses (i.e. Chlorophyll, Proteins, Carbohydrates, Lipids), the process of obtaining concentration values from plate-reading assays is also outlined in this code.    

We also used a RADseq pipeline in an attempt to identify corals that initially had unknown treatment identities (i.e. which site they were initially transplanted from). That pipeline can be found in the Github directory as well. Here we simply re-incorporated the three *S. siderea* corals that identified with high confidence using that pipeline by changing the "NA" listed in their home/collection site column in the initial/raw datasheets (`FvFm_All_corals.csv`, `mort_RT.csv`, `sample_treatments.csv`).  

# Photosynthetic efficiency (PAM)

## Load data and organize

```{r}
#| eval: false 

# navigate to directory with the data
#setwd("~/Raw_data/Field Data/PAM")

#Load and format code

FvFm_data_All_corals <- read_csv("Raw_data/Field Data/PAM/FvFm_All_corals.csv", 
                            col_types = cols(Fo_1 = col_number(), 
                                             Fm_1 = col_number(), Fv_1 = col_number(), 
                                             FvFm1 = col_number(), Fo_2 = col_number(), 
                                             Fm_2 = col_number(), Fv_2 = col_number(), 
                                             FvFm2 = col_number(), Fo_3 = col_number(), 
                                             Fm_3 = col_number(), Fv_3 = col_number(), 
                                             FvFm3 = col_number()))
#FvFm_data_All_corals <- read_csv("~/Desktop/Desktop - Callie’s MacBook Pro/Undergrad Research /Honors Thesis/Field Data/PAM/FvFm_All_corals.csv", 
#                            col_types = cols(Fo_1 = col_number(), 
#                                             Fm_1 = col_number(), Fv_1 = col_number(), 
#                                             FvFm1 = col_number(), Fo_2 = col_number(), 
#                                             Fm_2 = col_number(), Fv_2 = col_number(), 
#                                             FvFm2 = col_number(), Fo_3 = col_number(), 
#                                             Fm_3 = col_number(), Fv_3 = col_number(), 
#                                             FvFm3 = col_number()))
#View(FvFm_data_All_corals)

# reassign treatment name via new naming scheme 
FvFm_data_All_corals$treatment <-
  ifelse(FvFm_data_All_corals$Collection_site == "NS" & FvFm_data_All_corals$Transplant_site == "NS", "NS Native",
         ifelse(FvFm_data_All_corals$Collection_site == "NS" & FvFm_data_All_corals$Transplant_site == "OS", "Transplant to OS",
                ifelse(FvFm_data_All_corals$Collection_site == "OS" & FvFm_data_All_corals$Transplant_site == "OS", "OS Native",
                       ifelse(FvFm_data_All_corals$Collection_site == "OS" & FvFm_data_All_corals$Transplant_site == "NS", "Transplant to NS", NA))))


FvFm_all_corals = subset(FvFm_data_All_corals, select = -c(Fo_1,Fm_1,Fv_1,Fo_2,Fm_2,Fv_2,Fo_3,Fm_3,Fv_3))

long_FvFm_all_corals <- FvFm_all_corals %>%
  pivot_longer(cols = 'FvFm1':'FvFm3', names_to = "replicate", values_to = "FvFm")

#Prepare data for graphing
#=====Raw data=====
  #removing corals missing treatment ID
FvFm_all_corals_no_NAs = subset(long_FvFm_all_corals, treatment %in% c("NS Native","Transplant to NS", "OS Native", "Transplant to OS"))

  #adding differentiation of environments for graphing panes  
FvFm_all_corals_no_NAs<- FvFm_all_corals_no_NAs %>% 
  mutate(transplant_envir = ifelse(grepl("NS", treatment), "Nearshore", "Offshore"))
FvFm_all_corals_no_NAs<- FvFm_all_corals_no_NAs %>% 
  mutate(home_envir = ifelse(treatment == "NS Native" | treatment == "Transplant to OS", 
                             "Nearshore coral", "Offshore coral"))
#View(FvFm_all_corals_no_NAs)


  #calculate an average FvFm value for each coral (one point/row for each coral)
FvFm_all_coral_one_point_no_NAs <- FvFm_all_corals_no_NAs |> 
  group_by(species, flag_ID, species_code) |> 
  # mutate keeps all 3 rows but adds the average to all of them
  mutate(avg_FvFm = mean(FvFm, na.rm = TRUE)) |>
  ungroup() |>
  # Pivot wider, spreading both individual replicates and the average
  pivot_wider(names_from = "replicate", values_from = "FvFm")


#Save raw data for records + statistical analysis 
write_csv(FvFm_all_coral_one_point_no_NAs, file="Cleaned_data/FvFm_noNA_data.csv")
#write_xlsx(FvFm_all_coral_one_point_no_NAs, path="~/Desktop/Desktop - Callie’s MacBook Pro/Undergrad Research /Honors Thesis/Publication/Cleaned_data/FvFm_noNA_data.xlsx")

#=======Mean data (one point for each transplant treatment)=========
PAM_means_all_coral<-FvFm_all_coral_one_point_no_NAs %>%
  #select(!Time) %>%
  group_by(species, treatment) %>%
  drop_na(avg_FvFm) %>%
  summarize(meanFvFm=mean(avg_FvFm), sdFvFm=sd(avg_FvFm), n=n(), SEFvFm=(sdFvFm)/(sqrt(n)))
#View(PAM_means_all_coral)

  #removing nas
PAM_means_no_NAs = subset(PAM_means_all_coral, treatment %in% c("NS Native","Transplant to NS", "OS Native", "Transplant to OS"))
   #adding differentiation of environment assignments for panels in graph
PAM_means_no_NAs<- PAM_means_no_NAs %>% 
  mutate(transplant_envir = ifelse(grepl("NS", treatment), "Nearshore", "Offshore"))
PAM_means_no_NAs<- PAM_means_no_NAs %>% 
  mutate(home_envir = ifelse(treatment == "NS Native" | treatment == "Transplant to OS", 
                             "Nearshore coral", "Offshore coral"))

```

## Figure creation

```{r}
#Graph plot of Fv/Fm by species and transplant treatment 
pd = position_dodge(width = 0.2)
pjd = position_jitterdodge(jitter.width = NULL, dodge.width = 0.2) #CBB: Updated this to jitter just to see points better

  # Trying and failing to make species a factor to change the order of my species in the legend
#PAM_means_no_NAs$species_factor <- factor(PAM_means_no_NAs$species, levels = c("SSID", "PSTR"))
#PAM_means_no_NAs$species_relevel <- relevel(PAM_means_no_NAs$species, ref = "SSID")
#levels(PAM_means_no_NAs$species) #returns: "SSID" "PSTR" 

## CBB edit: what you were trying above:
PAM_means_no_NAs$species <- factor(PAM_means_no_NAs$species, levels = c("SSID","PSTR"))
FvFm_all_coral_one_point_no_NAs$species <- factor(FvFm_all_coral_one_point_no_NAs$species, levels = c("SSID","PSTR"))
levels(PAM_means_no_NAs$species) # just lets you check the levels

#PAM_means_no_NAs |> 
  #mutate(species = factor(), levels = c('S. siderea', 'P. strigosa')) |> 

PAM_Fig <- ggplot(data = PAM_means_no_NAs %>% drop_na(), aes(x= factor(treatment, levels = c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=meanFvFm, color = species, group=species)) +
  geom_point(size=3, position=pd) +
  geom_point(data = FvFm_all_coral_one_point_no_NAs, aes(x=treatment, y=avg_FvFm), position=pjd, size=2.5, alpha=0.3) +
  geom_line(position=pd) +
  geom_errorbar(aes(x=treatment, ymin=meanFvFm-SEFvFm, ymax=meanFvFm+SEFvFm), 
                width=0.2, linewidth=1, position=pd) +
  facet_wrap(~home_envir, scales='free_x') + #THIS LINE ISN'T WORKING
  theme_bw() + 
  theme(
        axis.title = element_text(size = 15), # make axes title text larger
        axis.text = element_text(size = 12), # make axes text larger  
        legend.title = element_text(size = 13), # make legend title larger 
        legend.text = element_text(size = 13, face = "italic"), # make legend text larger
        strip.text = element_text(size = 14)) +  # make text in facet title larger 
  guides(color = guide_legend(title = "Species")) + 
  xlab("Treatment") +
  ylab("Fv/Fm values") +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c('S. siderea', 'P. strigosa')) 

PAM_Fig2 <- ggplot(data = PAM_means_no_NAs %>% drop_na(), aes(x = transplant_envir, y=meanFvFm, color = home_envir, group = home_envir)) +
  geom_errorbar(aes(x=transplant_envir, ymin=meanFvFm-SEFvFm, ymax=meanFvFm+SEFvFm), width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = FvFm_all_coral_one_point_no_NAs, aes(x=transplant_envir, y=avg_FvFm), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~species, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c("NS", "OS")) +
  scale_x_discrete(labels = c("NS", "OS")) +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    #legend.title = element_blank(), # remove legend title 
    strip.text = element_text(size = 14, face = "italic")) +  # make text in facet title larger & italic
  guides(color = guide_legend(title = "Native Environment")) + 
  xlab("Transplant Environment") +
  ylab("Fv/Fm values")


# export graph as .jpg
ggsave("PAM_Fig_Sep_17_26.jpg", plot=PAM_Fig2, path="Final_figures", width = 10, height = 5, units = "in", dpi = 600)
```


## Statistical analysis

```{r}
#| eval: false

#Load cleaned data 
FvFm_data <- read_csv("Cleaned_data/FvFm_noNA_data.csv", 
                      col_types = cols(average_FvFm = col_number()))
#View(FvFm_data)

#Test for normality
  # Using a Shapiro-wilk test, our null hypothesis is that the sample data is not significantly different from a normal distribution.
    # p-value < 0.05 - reject the null hypothesis (should not assume the data is normally distributed)
    # p-value > 0.05 - fail to reject the null hypothesis (can assume the data is normally distributed)

#S. siderea by treatment
SSID_PAM_df<-subset(FvFm_data, species == "SSID")

shapiro.test(subset(SSID_PAM_df, treatment == "Transplant to NS")$avg_FvFm)
  # p-value > 0.05
#no SSID transplant to OS corals
shapiro.test(subset(SSID_PAM_df, treatment == "NS Native")$avg_FvFm)
  # p-value < 0.05 (reject the null hypothesis)
shapiro.test(subset(SSID_PAM_df, treatment == "OS Native")$avg_FvFm)
  # p-value > 0.05

#P. strigosa by treatment 
PSTR_PAM_df<-subset(FvFm_data, species == "PSTR")

#one PSTR transplant to NS coral
shapiro.test(subset(PSTR_PAM_df, treatment == "Transplant to OS")$avg_FvFm)
  # p-value > 0.05
shapiro.test(subset(PSTR_PAM_df, treatment == "NS Native")$avg_FvFm)
  # p-value < 0.05 (reject the null hypothesis)
shapiro.test(subset(PSTR_PAM_df, treatment == "OS Native")$avg_FvFm)
  # p-value > 0.05

#Test for equal variance
  # Null hypothesis: all group variances are equal
  # High p-value (p > 0.05): fail to reject the null (can assume that the variances are equal enough) 
  # Low p-value (p < 0.05): reject the null (should not assume that variances are equal enough)

leveneTest(avg_FvFm ~ treatment, SSID_PAM_df)
  # p-value > 0.05
leveneTest(avg_FvFm ~ treatment, PSTR_PAM_df)
  # p-value > 0.05

#Run ANOVAs + Tukey HSD
#======comparing species=======#
aovPAM<-aov(avg_FvFm~species, data=FvFm_data)
summary(aovPAM)
  # F = 12.88
  # p-value < 0.001
TukeyHSD(aovPAM)

#======comparing treatments for each species=======#
# S. siderea
aovPAMs<-aov(avg_FvFm~treatment, data=SSID_PAM_df)
summary(aovPAMs)

# P. strigosa
aovPAMp<-aov(avg_FvFm~treatment, data=PSTR_PAM_df)
summary(aovPAMp)
TukeyHSD(aovPAMp)


#Summary statistics 
stats_FvFm <-FvFm_data %>%
  group_by(species, treatment) %>%
  summarise(
          count = n(),
          mean_FvFm = mean(avg_FvFm))
stats_FvFm
```

```{r CBB stats run for PAM}

# just renaming to make it shorter
pam_df <- FvFm_all_coral_one_point_no_NAs %>% 
  mutate(across(c(2:5, 8:9, 11:13), as.factor)) # convert several columns to factors


# ------------- SSID-only PAM stats
## going to move forward with a model per species so separating into two different dfs starting with SSID:
SSID_pam <- subset(pam_df, species == "SSID")

# visual inspection of distribution (density + histogram): skewed
ggplot(SSID_pam, aes(x = avg_FvFm)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15) +
  geom_density()

# visual inspection of Q-Q plot: some outliers on extremes
ggplot(SSID_pam, aes(sample = avg_FvFm)) +
  stat_qq() +
  stat_qq_line(color = "coral")



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
SSID_pam_lm <- lmer(avg_FvFm ~ transplant_envir * home_envir + (1 | colony_ID), data = SSID_pam)
check_model(SSID_pam_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
SSID_pam_lm_trt <- lmer(avg_FvFm ~ treatment + (1 | colony_ID), data = SSID_pam)
compare_performance(SSID_pam_lm, SSID_pam_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for simple but RE if more biologically appropriate and does not significantly change interpretation
SSID_pam_lm_simple <- lm(avg_FvFm ~ treatment, data = SSID_pam)
AIC(SSID_pam_lm, SSID_pam_lm_simple)


# Stats output for selected model
summary(SSID_pam_lm_trt)
anova(SSID_pam_lm_trt) 
emmeans(SSID_pam_lm_trt, pairwise ~ treatment, adjust = "tukey")

## Final SSID PAM model: SSID_pam_lm_trt <- lmer(avg_FvFm ~ treatment + (1 | colony_ID), data = SSID_pam)



# ------------- PSTR-only PAM stats

PSTR_pam_full <- subset(pam_df, species != "SSID")
table(PSTR_pam_full$treatment) # only n=1 for 'Transplant to NS' so cutting that level

PSTR_pam <- PSTR_pam_full %>% 
  filter(treatment != "Transplant to NS") %>% # remove the treatment with n=1
  filter(flag_ID != "37") %>% # remove sample with bad PAM (explain in methods)
  droplevels()

# visual inspection of distribution (density + histogram): skewed
ggplot(PSTR_pam, aes(x = avg_FvFm)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15) +
  geom_density()

# visual inspection of Q-Q plot: couple outliers on extremes
ggplot(PSTR_pam, aes(sample = avg_FvFm)) +
  stat_qq() +
  stat_qq_line(color = "coral") 



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
PSTR_pam_lm <- lmer(avg_FvFm ~ transplant_envir * home_envir + (1 | colony_ID), data = PSTR_pam)
check_model(PSTR_pam_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
PSTR_pam_lm_trt <- lmer(avg_FvFm ~ treatment + (1 | colony_ID), data = PSTR_pam)
compare_performance(PSTR_pam_lm, PSTR_pam_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for simple but RE if more biologically appropriate and does not significantly change interpretation
PSTR_pam_lm_simple <- lm(avg_FvFm ~ treatment, data = PSTR_pam)
AIC(PSTR_pam_lm, PSTR_pam_lm_simple)


# Stats output for selected model
summary(PSTR_pam_lm_trt)
anova(PSTR_pam_lm_trt) 
emmeans(PSTR_pam_lm_trt, pairwise ~ treatment, adjust = "tukey")

## Final PSTR PAM model: PSTR_pam_lm_trt <- lmer(avg_FvFm ~ treatment + (1 | colony_ID), data = PSTR_pam)


# ------------ Followed this before subsetting data:
##### Performed all the below with full PSTR sample and from that determined we needed to remove 2 samples:
# the single Transplant to NS fragment
# the bad PAM sample

## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
#PSTR_pam_lm <- lmer(avg_FvFm ~ transplant_envir * home_envir + (1 | colony_ID), data = PSTR_pam)
# throw singular fit warnging...
#check_singularity(PSTR_pam_lm) # yes, singular so moving forward with add not interactive model
#PSTR_pam_lm <- lmer(avg_FvFm ~ transplant_envir + home_envir + (1 | colony_ID), data = PSTR_pam)
# still throwing singular fit do dropping random effect and shifting to simple lm

# PSTR_pam_lm <- lm((avg_FvFm) ~ transplant_envir * home_envir, data = PSTR_pam)
# check_model(PSTR_pam_lm) # use this to visually check model performance/assumptions: looks meh
# check_heteroscedasticity(PSTR_pam_lm) # OK: Error variance appears to be homoscedastic (p = 0.112).
# check_outliers(PSTR_pam_lm) # no corals considered outliers
# 
# summary(PSTR_pam_lm)
# anova(PSTR_pam_lm) 
# 
# ## Exploring difference between 2-way, 1-way, and additive: additive is best according to AIC so moving forward # with this one!
# PSTR_pam_lm_add <- lm(log(avg_FvFm) ~ transplant_envir + home_envir, data = PSTR_pam)
# PSTR_pam_lm_trt <- lm(log(avg_FvFm) ~ treatment, data = PSTR_pam)
# compare_performance(PSTR_pam_lm, PSTR_pam_lm_add, PSTR_pam_lm_trt)
# 
# check_model(PSTR_pam_lm_add) # use this to visually check model performance/assumptions: looks better
# shapiro.test(residuals(PSTR_pam_lm_add))
# # residual normality
# shapiro.test(residuals(PSTR_pam_lm_add)) 
# # W = 0.68253, p-value = 1.629e-06 --> very non-normal so moving forward with removing bad PAM sample

```


# Chlorophyll concentrations

## Load data and organize 

```{r}
#| eval: false
#Load data and format code
#Chlorophyll data
chlorophyll_data <- read_csv("Raw_data/Lab Data/Chloropyll/chlorophyll_data.csv", 
                             col_types = cols(absorbance = col_character(), 
                                              absorbance_reading = col_number()))
chl_samples<-filter(chlorophyll_data, chl_source=='sample')
#Volume of tissue slurry
chl_starting_volume <- read_csv("Raw_data/Lab Data/Chloropyll/chl_starting_volume.csv", 
                                col_types = cols(total_symbiont_slurry_volume_ml = col_number()))
final_chl<-merge(chl_samples, chl_starting_volume, by="sample_ID")

#After compiling data sets, group by individual (average of 3 replicates for each) and find the mean absorbance values for tested wavelengths of 632 nm and 665 nm 
chl_abs_means<- final_chl %>%
  group_by(sample_ID, absorbance, plate_number) %>%
  summarise(meanabsor=mean(absorbance_reading), sd=sd(absorbance_reading), n=n(), se=sd/sqrt(n), 
            meanvol=mean(total_symbiont_slurry_volume_ml))

#Use the following equations to calculate chlorophyll concentration (which will be in ug/mL)
    #First account for the path length in absorbance values 
    #Again, the blank averages from each plate are...     
        #Plate 5 Blank average = 0.07475 # calculated from plate reader (in lab notes)
        #Plate 6 Blank average = 0.074909 # calculated from plate reader (in lab notes)
chl_abs_means$blank<-ifelse(chl_abs_means$plate_number==5, 0.07475, 0.074909)

#And the path lengths of each plate are...
        #Plate 5 path length = 0.6493 # reported from plate reader (in lab notes)
        #Plate 6 path length = 0.6445 # reported from plate reader (in lab notes)
chl_abs_means$pathlength<-ifelse(chl_abs_means$plate_number==5, 0.6493, 0.6445)

#Equation for adjusting with path length is: absor @ given wavelength = (absor-blank)/pathlength (from Warren 2008)
chl_abs_means$adjusted_abs_value<-(chl_abs_means$meanabsor - chl_abs_means$blank)/chl_abs_means$pathlength
    #Then calculate concentrations using these equations from Ritchie (2006):
    #chla = -3.4551(A632) + 13.6849(A665)
    #chlc2 = 32.9371(A632) - 7.0140(A665)

#Let's make A632 and A665 values in their own column so that we can craft the necessary equation!
chl_abs_means_wider<-chl_abs_means %>%
  select(-c(meanabsor, sd, se)) %>%
  pivot_wider(names_from= absorbance, values_from = adjusted_abs_value)
#View(chl_abs_means_wider)

#Okay time to get the calculations in!
chl_abs_means_wider$chla_ugml<-((-3.4551*(chl_abs_means_wider$"632")) + (13.6849*(chl_abs_means_wider$"665")))
chl_abs_means_wider$chlc2_ugml<-((32.9371*(chl_abs_means_wider$"632")) - (7.0140*(chl_abs_means_wider$"665")))

  #(3) accommodate volumes of total tissue slurry bc want a count of ALL chlorophyll in the sample -- divide everything by 0.3846 (dilution factor) and multiply by 5 (for total mls)
chl_abs_means_wider$chla_ug<-(chl_abs_means_wider$chla_ugml * 5 / 0.3846)
chl_abs_means_wider$chlc2_ug<-(chl_abs_means_wider$chlc2_ugml * 5 / 0.3846)
#View(chl_abs_means_wider)

#Prepare data for graphing
#=====Raw data======
sa_measurements<-read_csv("Raw_data/Lab Data/backcalcs/sa_measurements.csv")
chl_sa_analysis<-merge(chl_abs_means_wider, sa_measurements, by="sample_ID")

#back-calculate each chlorophyll concentration (a and c2) to ug/cm^2 and add coral treatments
chl_sa_analysis$chla_ug_cm2<- chl_sa_analysis$chla_ug / chl_sa_analysis$sample_area_cm2
chl_sa_analysis$chlc2_ug_cm2<- chl_sa_analysis$chlc2_ug / chl_sa_analysis$sample_area_cm2

samples_treatments<-read_csv("Raw_data/Lab Data/backcalcs/samples_treatments.csv")
chl_ugcm2_treatments<-merge(chl_sa_analysis, samples_treatments, by="sample_ID") #RERUNING THIS LINE SHOULD ACCOUNT FOR TREATMENTS FOR SSIDF3 and SSIDF18 

  # Fix treatment naming convention
chl_ugcm2_treatments$treatment<-ifelse(chl_ugcm2_treatments$treatment == 'Native NS', 'NS Native', 
                    as.character(chl_ugcm2_treatments$treatment))
chl_ugcm2_treatments$treatment<-ifelse(chl_ugcm2_treatments$treatment == 'Native OS', 'OS Native', 
                    as.character(chl_ugcm2_treatments$treatment))
  

  #save chl concentration data for records, graphing, and statistic
write_csv(chl_ugcm2_treatments, file="Cleaned_data/chl_data.csv")
#write_xlsx(chl_ugcm2_treatments, path="~/Cleaned_data/chl_data.xlsx")

```

## Figure creation for chlorophyll-a and chlorophyll-c2

```{r}
  #removing corals missing treatment ID
chl_ugcm2_treatments_noNA<-filter(chl_ugcm2_treatments, treatment!="na")

  #changing two chl-a values < 0 to equal 0 based on literature
chl_ugcm2_treatments_noNA$chla_ug_cm2 = ifelse(chl_ugcm2_treatments_noNA$chla_ug_cm2 < 0, 0, chl_ugcm2_treatments_noNA$chla_ug_cm2)

   #adding differentiation of environment assignments for panels in graph
chl_ugcm2_treatments_noNA<- chl_ugcm2_treatments_noNA %>% 
  mutate(transplant_envir = ifelse(grepl("NS", treatment), "Nearshore", "Offshore"))
chl_ugcm2_treatments_noNA<- chl_ugcm2_treatments_noNA %>% 
  mutate(home_envir = ifelse(treatment == "NS Native" | treatment == "Transplant to OS", 
                             "Nearshore coral", "Offshore coral"))

#Mean data
Chl_mean_plot_data <- chl_ugcm2_treatments_noNA %>%
  group_by(species, treatment, transplant_envir, home_envir) %>%
  summarise(# clh-a
            meanchla=mean(chla_ug_cm2), sdchla=sd(chla_ug_cm2), 
            nchla=n(), sechla=sdchla/sqrt(nchla), 
            # chl-c2
            meanchlc2=mean(chlc2_ug_cm2), sdchlc2=sd(chlc2_ug_cm2), 
            nchlc2=n(), sechlc2=sdchlc2/sqrt(nchlc2))

## CBB edit: what you were trying above:
Chl_mean_plot_data$species <- factor(Chl_mean_plot_data$species, levels = c("SSID","PSTR"))
chl_ugcm2_treatments_noNA$species <- factor(chl_ugcm2_treatments_noNA$species, levels = c("SSID","PSTR"))

#======Begin graphing======

#Graph plot of chlorophyll-a by species and transplant treatment 

chla_Fig <- ggplot(data = Chl_mean_plot_data %>% drop_na(), aes(x= factor(treatment, levels = c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=meanchla, color = species, group=species)) +
  geom_errorbar(aes(x=treatment, ymin=meanchla-sechla, ymax=meanchla+sechla), 
                width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = chl_ugcm2_treatments_noNA, aes(x=treatment, y=chla_ug_cm2), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~home_envir, scales='free_x') +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    legend.title = element_text(size = 13), # make legend title larger 
    legend.text = element_text(size = 13, face = "italic"), # make legend text larger
    strip.text = element_text(size = 14)) +  # make text in facet title larger 
  guides(color = guide_legend(title = "Species")) + 
  xlab("Treatment") +
  ylab("Mean [chl-a] (ug/cm^2)")+
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c('S. siderea', 'P. strigosa')) 

chla_Fig2 <- ggplot(data = Chl_mean_plot_data %>% drop_na(), aes(x = transplant_envir, y=meanchla, color = home_envir, group = home_envir)) +
  geom_errorbar(aes(x=transplant_envir, ymin=meanchla-sechla, ymax=meanchla+sechla), width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = chl_ugcm2_treatments_noNA, aes(x=transplant_envir, y=chla_ug_cm2), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~species, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c("NS", "OS")) +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    #legend.title = element_blank(), # remove legend title 
    strip.text = element_text(size = 14, face = "italic")) +  # make text in facet title larger & italic
  guides(color = guide_legend(title = "Native Environment")) + 
  xlab("Transplant Environment") +
  ylab("Mean [chl-a] (ug/cm^2)")


# export graph as .jpg
ggsave("Chla_Fig_Sep_17_26.jpg", plot=chla_Fig2, path="Final_figures", width = 10, height = 5, units = "in", dpi = 600)




#Graph plot of chlorophyll-c2 by species and transplant treatment
chlc2_Fig <- ggplot(data = Chl_mean_plot_data %>% drop_na(), aes(x= factor(treatment, levels = c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=meanchlc2, color = species, group=species)) +
  geom_errorbar(aes(x=treatment, ymin=meanchlc2-sechlc2, ymax=meanchlc2+sechlc2), 
                width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = chl_ugcm2_treatments_noNA, aes(x=treatment, y=chlc2_ug_cm2), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~home_envir, scales='free_x') +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    legend.title = element_text(size = 13), # make legend title larger 
    legend.text = element_text(size = 13, face = "italic"), # make legend text larger
    strip.text = element_text(size = 14)) +  # make text in facet title larger 
  guides(color = guide_legend(title = "Species")) + 
  xlab("Treatment") +
  ylab("Mean [chl-c2] (ug/cm^2)")+
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c('S. siderea', 'P. strigosa')) 

chlc2_Fig2 <- ggplot(data = Chl_mean_plot_data %>% drop_na(), aes(x = transplant_envir, y=meanchlc2, color = home_envir, group = home_envir)) +
  geom_errorbar(aes(x=transplant_envir, ymin=meanchlc2-sechlc2, ymax=meanchlc2+sechlc2), width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = chl_ugcm2_treatments_noNA, aes(x=transplant_envir, y=chlc2_ug_cm2), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~species, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c("Nearshore", "Offshore")) +
  scale_x_discrete(labels = c("NS", "OS")) +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    #legend.title = element_blank(), # remove legend title 
    strip.text = element_text(size = 14, face = "italic")) +  # make text in facet title larger & italic
  guides(color = guide_legend(title = "Home Environment")) + 
  xlab("Transplant Environment") +
  ylab("Mean [chl-c2] (ug/cm^2)")

# export graph as .jpg
ggsave("Chlc2_Fig_Sep_17_26.jpg", plot=chlc2_Fig2, path="Final_figures", width = 10, height = 5, units = "in", dpi = 600)

#Save graphed raw data (w/out NA in transplant treatment) for statistical analysis 
#write_csv(chl_ugcm2_treatments_noNA, file="Cleaned_data/chl_noNA_data.csv")
```


## Statistical analysis for chlorophyll-a

```{r}
#| eval: false
#Load data 
#chl_analysis_data <- read_csv("Cleaned_data/chl_noNA_data.csv", 
#    col_types = cols(chla_ug_cm2 = col_number(), 
#        chlc2_ug_cm2 = col_number()))

#Test for normality

#S. siderea by treatment
SSID_chl_cm2<-subset(chl_analysis_data, species == "SSID")

shapiro.test(subset(SSID_chl_cm2, treatment == "Transplant to NS")$chla_ug_cm2)
  # p-value > 0.05
#no SSID transplant to OS corals
shapiro.test(subset(SSID_chl_cm2, treatment == "OS Native")$chla_ug_cm2)
  # p-value > 0.05
shapiro.test(subset(SSID_chl_cm2, treatment == "NS Native")$chla_ug_cm2)
  # p-value > 0.05

#P. strigosa by treatment 
PSTR_chl_cm2<-subset(chl_analysis_data, species == "PSTR")

#one PSTR transplant to NS coral
shapiro.test(subset(PSTR_chl_cm2, treatment == "Transplant to OS")$chla_ug_cm2)
  # p-value > 0.05
shapiro.test(subset(PSTR_chl_cm2, treatment == "OS Native")$chla_ug_cm2)
  # p-value > 0.05
shapiro.test(subset(PSTR_chl_cm2, treatment == "NS Native")$chla_ug_cm2)
  # p-value > 0.05

#Test for equal variance
leveneTest(chla_ug_cm2 ~ treatment, SSID_chl_cm2)
  # p > 0.05
leveneTest(chla_ug_cm2 ~ treatment, PSTR_chl_cm2)
  # p > 0.05

#Run ANOVAs + Tukey HSD
#======comparing species=======#
aovchla<-aov(chla_ug_cm2~species, data = chl_analysis_data)
summary(aovchla)
#======comparing treatments for each species=======#
  #P. strigsosa
aovchlap<-aov(chla_ug_cm2~treatment, data = PSTR_chl_cm2)
summary(aovchlap)
TukeyHSD(aovchlap)
  #S. siderea
aovchlas<-aov(chla_ug_cm2~treatment, data = SSID_chl_cm2)
summary(aovchlas)

#Summary statistics 
stats_chla <-chl_analysis_data %>%
  group_by(species, treatment) %>%
  summarise(
          count = n(),
          mean_chla = mean(chla_ug_cm2))
stats_chla
```

```{r CBB chla stats}


# just renaming to make it shorter
chla_df <- chl_ugcm2_treatments_noNA %>% 
  mutate(across(c(17:20), as.factor)) %>% # convert several columns to factors
  mutate(colony_ID = str_extract(sample_ID, "[0-9]")) # need to add colony ID for random effects


# ------------- SSID-only Chla stats
## going to move forward with a model per species so separating into two different dfs starting with SSID:
SSID_chla <- subset(chla_df, species == "SSID")


# visual inspection of distribution (density + histogram): skewed
ggplot(SSID_chla, aes(x = chla_ug_cm2)) +
  geom_histogram(aes(y = after_stat(density)), bins = 10) +
  geom_density()

# visual inspection of Q-Q plot: some outliers on extremes
ggplot(SSID_chla, aes(sample = chla_ug_cm2)) +
  stat_qq() +
  stat_qq_line(color = "coral")



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
SSID_chla_lm <- lmer(chla_ug_cm2 ~ transplant_envir * home_envir + (1 | colony_ID), data = SSID_chla)
check_model(SSID_chla_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
SSID_chla_lm_trt <- lmer(chla_ug_cm2 ~ treatment + (1 | colony_ID), data = SSID_chla)
compare_performance(SSID_chla_lm, SSID_chla_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for simple but RE if more biologically appropriate and does not significantly change interpretation
SSID_chla_lm_simple <- lm(chla_ug_cm2 ~ treatment, data = SSID_chla)
AIC(SSID_chla_lm, SSID_chla_lm_simple)


# Stats output for selected model
summary(SSID_chla_lm_simple)
anova(SSID_chla_lm_simple) 
emmeans(SSID_chla_lm_simple, pairwise ~ treatment, adjust = "tukey")

## Final SSID chl a model: SSID_chla_lm_simple <- lm(chla_ug_cm2 ~ treatment, data = SSID_chla)



# ------------- PSTR-only Chla stats

PSTR_chla_full <- subset(chla_df, species != "SSID")
table(PSTR_chla_full$treatment) # only n=1 for 'Transplant to NS' so cutting that level

PSTR_chla <- PSTR_chla_full %>% 
  filter(treatment != "Transplant to NS") %>% # remove the treatment with n=1
  droplevels()

# visual inspection of distribution (density + histogram): skewed
ggplot(PSTR_chla, aes(x = chla_ug_cm2)) +
  geom_histogram(aes(y = after_stat(density)), bins = 13) +
  geom_density()

# visual inspection of Q-Q plot: couple outliers on extremes
ggplot(PSTR_chla, aes(sample = chla_ug_cm2)) +
  stat_qq() +
  stat_qq_line(color = "coral") 



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
PSTR_chla_lm <- lmer(chla_ug_cm2 ~ transplant_envir * home_envir + (1 | colony_ID), data = PSTR_chla)
check_model(PSTR_chla_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
PSTR_chla_lm_trt <- lmer(chla_ug_cm2 ~ treatment + (1 | colony_ID), data = PSTR_chla)
compare_performance(PSTR_chla_lm, PSTR_chla_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for simple but RE if more biologically appropriate and does not significantly change interpretation
PSTR_chla_lm_simple <- lm(chla_ug_cm2 ~ treatment, data = PSTR_chla)
AIC(PSTR_chla_lm, PSTR_chla_lm_simple)


# Stats output for selected model
summary(PSTR_chla_lm_simple)
anova(PSTR_chla_lm_simple) 
emmeans(PSTR_chla_lm_simple, pairwise ~ treatment, adjust = "tukey")

## Final PSTR chl a model: PSTR_chla_lm_simple <- lm(chla_ug_cm2 ~ treatment, data = PSTR_chla)

```


## Statistical analysis for chlorophyll-c2

```{r}
#| eval: false
# If needed, reload data
#Load data 
# chl_analysis_data <- read_csv("~/Cleaned_data/chl_noNA_data.csv", 
#     col_types = cols(chla_ug_cm2 = col_number(), 
#         chlc2_ug_cm2 = col_number()))

#Test for normality

#S. siderea by treatment
SSID_chl_cm2<-subset(chl_analysis_data, species == "SSID")

shapiro.test(subset(SSID_chl_cm2, treatment == "Transplant to NS")$chlc2_ug_cm2)
  # p-value > 0.05   
#no SSID transplant to OS corals
shapiro.test(subset(SSID_chl_cm2, treatment == "OS Native")$chlc2_ug_cm2)
  # p-value > 0.05 
shapiro.test(subset(SSID_chl_cm2, treatment == "NS Native")$chlc2_ug_cm2)
  # p-value > 0.05 

#P. strigosa by treatment 
PSTR_chl_cm2<-subset(chl_analysis_data, species == "PSTR")

#one PSTR transplant to NS coral
shapiro.test(subset(PSTR_chl_cm2, treatment == "Transplant to OS")$chlc2_ug_cm2)
  # p-value > 0.05 
shapiro.test(subset(PSTR_chl_cm2, treatment == "OS Native")$chlc2_ug_cm2)
  # p-value > 0.05 
shapiro.test(subset(PSTR_chl_cm2, treatment == "NS Native")$chlc2_ug_cm2)
  # p-value > 0.05 

#Test for equal variance
leveneTest(chlc2_ug_cm2 ~ treatment, SSID_chl_cm2)
  # p-value > 0.05 
leveneTest(chlc2_ug_cm2 ~ treatment, PSTR_chl_cm2)
  # p-value > 0.05 

#Run ANOVAs + Tukey HSD
#======comparing species=======#
aovchlc2<-aov(chlc2_ug_cm2~species, data = chl_analysis_data)
summary(aovchlc2)
#======comparing treatments for each species=======#
  #P. strigsosa
aovchlc2p<-aov(chlc2_ug_cm2~treatment, data = PSTR_chl_cm2)
summary(aovchlc2p)
TukeyHSD(aovchlc2p)
  #S. siderea
aovchlc2s<-aov(chlc2_ug_cm2~treatment, data = SSID_chl_cm2)
summary(aovchlc2s)

#Summary statistics
stats_chlc2 <-chl_analysis_data %>%
  group_by(species, treatment) %>%
  summarise(
          count = n(),
          mean_chlc2 = mean(chlc2_ug_cm2))
stats_chlc2
```

```{r CBB chlc2 stats}


# ------------- SSID-only Chlc2 stats
# visual inspection of distribution (density + histogram): skewed
ggplot(SSID_chla, aes(x = chlc2_ug_cm2)) +
  geom_histogram(aes(y = after_stat(density)), bins = 12) +
  geom_density()

# visual inspection of Q-Q plot: some outliers on extremes
ggplot(SSID_chla, aes(sample = chlc2_ug_cm2)) +
  stat_qq() +
  stat_qq_line(color = "coral")



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
SSID_chlc2_lm <- lmer(chlc2_ug_cm2 ~ transplant_envir * home_envir + (1 | colony_ID), data = SSID_chla)
check_model(SSID_chlc2_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
SSID_chlc2_lm_trt <- lmer(chlc2_ug_cm2 ~ treatment + (1 | colony_ID), data = SSID_chla)
compare_performance(SSID_chlc2_lm, SSID_chlc2_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for simple but RE if more biologically appropriate and does not significantly change interpretation
SSID_chlc2_lm_simple <- lm(chlc2_ug_cm2 ~ treatment, data = SSID_chla)
AIC(SSID_chlc2_lm, SSID_chlc2_lm_simple)


# Stats output for selected model
summary(SSID_chlc2_lm_simple)
anova(SSID_chlc2_lm_simple) 
emmeans(SSID_chlc2_lm_simple, pairwise ~ treatment, adjust = "tukey")

## Final SSID chl c2 model: SSID_chlc2_lm_simple <- lm(chlc2_ug_cm2 ~ treatment, data = SSID_chla)



# ------------- PSTR-only Chla stats

# visual inspection of distribution (density + histogram): skewed
ggplot(PSTR_chla, aes(x = chlc2_ug_cm2)) +
  geom_histogram(aes(y = after_stat(density)), bins = 6) +
  geom_density()

# visual inspection of Q-Q plot: couple outliers on extremes
ggplot(PSTR_chla, aes(sample = chlc2_ug_cm2)) +
  stat_qq() +
  stat_qq_line(color = "coral") 



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
PSTR_chlc2_lm <- lmer(chlc2_ug_cm2 ~ transplant_envir * home_envir + (1 | colony_ID), data = PSTR_chla)
check_model(PSTR_chlc2_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
PSTR_chlc2_lm_trt <- lmer(chlc2_ug_cm2 ~ treatment + (1 | colony_ID), data = PSTR_chla)
compare_performance(PSTR_chlc2_lm, PSTR_chlc2_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for simple but RE if more biologically appropriate and does not significantly change interpretation
PSTR_chlc2_lm_simple <- lm(chlc2_ug_cm2 ~ treatment, data = PSTR_chla)
AIC(PSTR_chlc2_lm, PSTR_chlc2_lm_simple)


# Stats output for selected model
summary(PSTR_chlc2_lm_simple)
anova(PSTR_chlc2_lm_simple) 
emmeans(PSTR_chlc2_lm_simple, pairwise ~ treatment, adjust = "tukey")

## Final PSTR chl c2 model: PSTR_chlc2_lm_simple <- lm(chlc2_ug_cm2 ~ treatment, data = PSTR_chla)

```

# Mean + Percent Change in Buoyant weight

## **NOT SURE HOW TO FIX WITH RE-IDENTIFIED SSID SAMPLES -- don't we need to know who specifically they are?** ##

## Load data and organize

```{r}
#| eval: false
#Load and format code
all_corals_and_TPs <- read_csv("Raw_data/Field Data/BW/all_corals_and_TPs.csv", 
                               col_types = cols(mortality = col_character(), 
                                                bw = col_number(), bw_sd = col_number(), 
                                                bw_se = col_number()))
#View(all_corals_and_TPs)
all_corals_and_TPs <- all_corals_and_TPs[,-7]

  #adding differentiation of transplant treatment
all_corals_and_TPs$treatment <- ifelse(all_corals_and_TPs$collection_site == "NS" & all_corals_and_TPs$transplant_site == "NS", "NS Native", ifelse(all_corals_and_TPs$collection_site == "NS" & all_corals_and_TPs$transplant_site == "OS", "Transplant to OS",
                ifelse(all_corals_and_TPs$collection_site == "OS" & all_corals_and_TPs$transplant_site == "OS", "OS Native",
                       ifelse(all_corals_and_TPs$collection_site == "OS" & all_corals_and_TPs$transplant_site == "NS", "Transplant to NS", NA))))

#Prepare data for graphing mean buoyant weight figure 
#======Raw data=======
alive_corals_and_TPs <- subset(all_corals_and_TPs, mortality %in% c('alive'))

#======Mean data=======
BW_alive_TP_means<-alive_corals_and_TPs %>%
  group_by(species, treatment, timepoint) %>%
  drop_na(treatment) %>%
  summarize(meanBW=mean(bw), sdBW=sd(bw), n=n(), SEBW=(sdBW)/(sqrt(n)))
#View(BW_alive_TP_means)
```

## Figure creation

```{r}
#Graph plot of % change in buoyant weight by species, transplant treatment, and time point 
pd_bw = position_dodge(width = 0.7)
new_species <- c("PSTR" = "P. strigosa", "SSID" = "S. siderea")

BW_alive_TP_plot <- ggplot(data= BW_alive_TP_means[!is.na(BW_alive_TP_means$treatment), ], 
      aes(x=factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=meanBW, color=timepoint)) +
  geom_point(data = alive_corals_and_TPs[!is.na(alive_corals_and_TPs$treatment), ], 
             aes(x=factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=bw), position=pd_bw, size=2, alpha=0.2) +
  geom_errorbar(aes(ymin=meanBW-SEBW, ymax=meanBW+SEBW), position=pd_bw, width=0.3) +
  geom_point(size=2, position=pd_bw) +
  xlab("Treatment") +
  ylab("Mean buoyant weight (g)") +
  guides(color = guide_legend(title = "Time point")) + 
  scale_color_manual(values=c("#b22222", "#228b21", "#8b4500", "#1f8fff","#7a5eb2")) +
  facet_wrap(~species, ncol=1, labeller = labeller(species = new_species)) +
  theme_bw()
BW_alive_TP_plot +
  theme(strip.text.x = element_text(face="italic"))

#Prepare data for graphing percent change buoyant weight figure 
wide_alive_coralsTPs <- alive_corals_and_TPs %>%
  pivot_wider(-bw_date, names_from = timepoint, values_from = c(bw, bw_sd, bw_se))

wide_alive_coralsTPs$T0_T1 <- with(wide_alive_coralsTPs, (bw_T1 - bw_T0)/(bw_T0))
wide_alive_coralsTPs$T0_T2 <- with(wide_alive_coralsTPs, (bw_T2 - bw_T0)/(bw_T0))
wide_alive_coralsTPs$T0_T3 <- with(wide_alive_coralsTPs, (bw_T3 - bw_T0)/(bw_T0))
wide_alive_coralsTPs$T0_T4 <- with(wide_alive_coralsTPs, (bw_T4 - bw_T0)/(bw_T0))
wide_alive_coralsTPs$T3_T4 <- with(wide_alive_coralsTPs, (bw_T4 - bw_T3)/(bw_T3))

long_alive_corals <- wide_alive_coralsTPs %>%
  pivot_longer(cols = c('T0_T1', 'T0_T2', 'T0_T3', 'T0_T4', 'T3_T4'), names_to ='timepoint_comparison', values_to = 'percent_change_bw')

#Raw data
long_alive_corals2<-long_alive_corals |>
  drop_na(percent_change_bw)

#Mean data
BW_timepointcomparison_means <-long_alive_corals %>%
  group_by(species, treatment, timepoint_comparison) %>%
  drop_na(percent_change_bw) %>%
  summarize(meanchange=mean(percent_change_bw), sdchange=sd(percent_change_bw), n=n(), SEchange=(sdchange)/(sqrt(n)))
#View(BW_timepointcomparison_means)

#Graph plot of percent change in buoyant weight by species and transplant treatment 
ggplot(data=BW_timepointcomparison_means,  
                    aes(x=factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=meanchange, color=timepoint_comparison)) +
  geom_point(data=long_alive_corals2, aes(x=factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=percent_change_bw, color = timepoint_comparison), position=pd, size=1, alpha=0.2) +
  geom_point(size=2, position=pd) +
  geom_errorbar(aes(ymin=meanchange-SEchange, ymax=meanchange+SEchange, group=timepoint_comparison), position=pd, width=0.4, color='black') +
  facet_wrap(~species, ncol=1, labeller = labeller(species = new_species), scales='free') +
  theme_bw() +
  theme(strip.text.x = element_text(face="italic")) + 
  xlab("Treatment") +
  ylab("Mean % change in buoyant weight (g)") + 
  scale_color_manual(name = "Timepoint comparison", labels=c('T0 - T1', 'T0 - T2', 'T0 - T3', 'T0 - T4', 'T3 - T4'), values=c("#b22222", "#228b21", "#8b4500", "#1f8fff","#7a5eb2"))
  #ylim(0,11)+

#Save raw data for statistical analysis --> CBB: I cut this since you can just reread it below rather than save it as its own csv
#write_csv(long_alive_corals2, file="long_alive_corals")
longcoral <- long_alive_corals2
```


### 2/24/26 Updated % change BW figure for BEM poster 
% change BW FIGURE modified for BEM Post Presentation 

Modifications made: 
- get rid of grams on y
- make points larger
- make text larger
- make border for each sub plot match PAM figure 
- change timepoints to months (T0 - T1) to (0 - 3)

```{r}

BW_Change_Fig <- ggplot(data=BW_timepointcomparison_means, aes(x=factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), 
                                              y=meanchange, color=timepoint_comparison)) + #mean data
  geom_point(data=long_alive_corals2, 
             aes(x=factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), 
                 y=percent_change_bw, 
                 color = timepoint_comparison), 
             position=pd, size=2, alpha=0.3) + #raw data 
  geom_point(size=3, position=pd) +
  geom_errorbar(aes(ymin=meanchange-SEchange, 
                    ymax=meanchange+SEchange, 
                    group=timepoint_comparison), position=pd) +
  facet_wrap(~species, ncol=1, labeller = labeller(species = new_species), scales='fixed') +
  theme_bw() +
  theme(strip.text.x = element_text(face="italic", size = 12),
        axis.title = element_text(size = 15), # make axes title text larger
        axis.text.y = element_text(size = 12), # make axes text larger
        axis.text.x = element_text(size = 13), #angle = 10, vjust = 0.75),
        legend.title = element_text(size = 14), # make legend title larger 
        legend.text = element_text(size = 13)) + # make legend text larger
  xlab("Treatment") +
  ylab("Percent change in buoyant weight") + 
  scale_color_manual(name = "Months:", labels=c('0 - 3', '0 - 9', '0 - 17', '0 - 60', '17 - 60'), values=c("#b22222", "#228b21", "#8b4500", "#1f8fff","#7a5eb2"))


```

```{r CBB BW plot edit}


bw_mean_df <- BW_timepointcomparison_means %>% 
  mutate(treat = treatment) %>% 
  mutate(treat = gsub("NS Native", "NS_NS", treat),
         treat = gsub("OS Native", "OS_OS", treat),
         treat = gsub("Transplant to NS", "OS_NS", treat),
         treat = gsub("Transplant to OS", "NS_OS", treat)) %>% 
  separate(treat, c("nat_env", "tran_env"))


BW_Change_Fig2 <- ggplot(data = bw_mean_df %>% filter(timepoint_comparison == "T0_T3" | timepoint_comparison == "T0_T4" & n >= 2), aes(x = factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), y=meanchange, color = timepoint_comparison)) +
  geom_errorbar(aes(x=factor(treatment, level=c('NS Native', 'Transplant to OS', 'OS Native', 'Transplant to NS')), ymin=meanchange-SEchange, ymax=meanchange+SEchange), width=0.25, linewidth=0.7, position=pd) +
  geom_point(size=3, position=pd) +
  #geom_point(data = long_alive_corals2 %>% filter(timepoint_comparison == "T0_T4"), aes(x=treatment, y=percent_change_bw), position=pjd, size=2.5, alpha=0.2) +
  facet_wrap(~species, ncol = 1, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) + 
  scale_color_manual(name = "Months:", labels=c('0 - 17', '0 - 60'), values=c("#8b4500", "#1f8fff")) +
  theme_bw() + 
  theme(strip.text.x = element_text(face="italic", size = 12),
        axis.title = element_text(size = 15), # make axes title text larger
        axis.text.y = element_text(size = 12), # make axes text larger
        axis.text.x = element_text(size = 13), #angle = 10, vjust = 0.75),
        legend.title = element_text(size = 14), # make legend title larger 
        legend.text = element_text(size = 13)) + # make legend text larger
  xlab("Transplant Environment") +
  ylab("Percent change in buoyant weight")







# bw to match other phys figures
BW_Change_Fig3 <- ggplot(data = bw_mean_df %>% filter(timepoint_comparison == "T0_T4" & n >= 2), aes(x = tran_env, y = meanchange, color = nat_env, group = nat_env)) +
  geom_errorbar(aes(x=tran_env, ymin=meanchange-SEchange, ymax=meanchange+SEchange), width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = long_alive_corals2 %>% filter(timepoint_comparison == "T0_T4"), aes(x=transplant_site, y=percent_change_bw, color = collection_site, group = collection_site), position=pjd, size=2.5, alpha=0.2) +
  facet_wrap(~species, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c("Nearshore", "Offshore")) +
  scale_x_discrete(labels = c("NS", "OS")) +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    #legend.title = element_blank(), # remove legend title 
    strip.text = element_text(size = 14, face = "italic")) +  # make text in facet title larger & italic
  guides(color = guide_legend(title = "Native Environment")) + 
  xlab("Transplant Environment") +
  ylab("Percent change in buoyant weight")
  


# export graph as .jpg 
ggsave("BWChange_Fig_Sep_17_26.jpg", plot=BW_Change_Fig3, path="Final_figures", width = 10, height = 5, units = "in", dpi = 600)

```


## Statistical analysis

```{r}
#| eval: false
#Load data 
#longcoral<-read_csv('long_alive_corals.csv')
longcoral$species<-ifelse(grepl("PSTR", longcoral$coral_ID), "PSTR", "SSID")

#Test for normality

#S. siderea by treatment
SSID_bw<-subset(longcoral, species == "SSID")
shapiro.test(subset(SSID_bw, treatment == "Transplant to NS")$percent_change_bw)
#no SSID transplant to OS corals
shapiro.test(subset(SSID_bw, treatment == "OS Native")$percent_change_bw)
shapiro.test(subset(SSID_bw, treatment == "NS Native")$percent_change_bw)

#P. strigosa by treatment 
PSTR_bw<-subset(longcoral, species == "PSTR")
#one PSTR transplant to NS coral
shapiro.test(subset(PSTR_bw, treatment == "Transplant to OS")$percent_change_bw)
shapiro.test(subset(PSTR_bw, treatment == "OS Native")$percent_change_bw)
shapiro.test(subset(PSTR_bw, treatment == "NS Native")$percent_change_bw)

#Test for equal variance
leveneTest(percent_change_bw ~ treatment, SSID_bw)
leveneTest(percent_change_bw ~ treatment, PSTR_bw)

#Run ANOVAs + Tukey HSD
aovbw1<-aov(percent_change_bw~treatment*species*timepoint_comparison, data=longcoral)
summary(aovbw1)
bw_results<-TukeyHSD(aovbw1)

#export results to sort outside of R
tidy_bw<-tidy(bw_results)
#View(tidy_bw)
write.csv(tidy_bw, file = "bw_results.csv")

#Summary statistics 
stats_bw <-longcoral %>%
  group_by(treatment, species, timepoint_comparison) %>%
  summarise(
          count = n(),
          mean_perc_bw = mean(percent_change_bw))
```

```{r CBB bw stats}

# just renaming to make it shorter
bw_df <- longcoral %>% 
  mutate(colony_ID = factor(str_extract(ID, "[0-9]"))) # need to add colony ID for random effects


# ------------- SSID-only BW stats
## going to move forward with a model per species so separating into two different dfs starting with SSID:
SSID_bw <- subset(bw_df, species == "SSID") %>% 
  filter(timepoint_comparison == "T0_T4")


# visual inspection of distribution (density + histogram): skewed
ggplot(SSID_bw, aes(x = percent_change_bw)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15) +
  geom_density()

# visual inspection of Q-Q plot: some outliers on extremes
ggplot(SSID_bw, aes(sample = percent_change_bw)) +
  stat_qq() +
  stat_qq_line(color = "coral")



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
SSID_bw_lm <- lmer(percent_change_bw ~ transplant_site * collection_site + (1 | colony_ID), data = SSID_bw)
check_model(SSID_bw_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
SSID_bw_lm_trt <- lmer(percent_change_bw ~ treatment + (1 | colony_ID), data = SSID_bw)
compare_performance(SSID_bw_lm, SSID_bw_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for RE model and it is more biologically appropriate
SSID_bw_lm_simple <- lm(percent_change_bw ~ treatment, data = SSID_bw)
AIC(SSID_bw_lm, SSID_bw_lm_simple)


# Stats output for selected model
summary(SSID_bw_lm_trt)
anova(SSID_bw_lm_trt) 
emmeans(SSID_bw_lm_trt, pairwise ~ treatment, adjust = "tukey")

## Final SSID BW model: SSID_bw_lm_trt <- lmer(percent_change_bw ~ treatment + (1 | colony_ID), data = SSID_bw)



# ------------- PSTR-only BW stats

# selecing for PSTR only:
PSTR_bw <- subset(bw_df, species != "SSID") %>% 
  filter(timepoint_comparison == "T0_T4") %>% 
  filter(treatment != "Transplant to NS") # remove the treatment with n=1

# visual inspection of distribution (density + histogram): okay
ggplot(PSTR_bw, aes(x = percent_change_bw)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15) +
  geom_density()

# visual inspection of Q-Q plot: couple outliers on extremes
ggplot(PSTR_bw, aes(sample = percent_change_bw)) +
  stat_qq() +
  stat_qq_line(color = "coral") 



## Since coral fragments may come from the same colony, we are going to run LMER with random effect of colony
PSTR_bw_lm <- lmer(percent_change_bw ~ transplant_site * collection_site + (1 | colony_ID), data = PSTR_bw)
check_model(PSTR_bw_lm) # use this to visually check model performance/assumptions. Looks good for these data!

## Exploring difference between 2-way and 1-way: no difference so proceeding with the one-way since we have to drop a treatment level
PSTR_bw_lm_trt <- lmer(percent_change_bw ~ treatment + (1 | colony_ID), data = PSTR_bw)
compare_performance(PSTR_bw_lm, PSTR_bw_lm_trt)

## Exploring simple vs random effect model: 
# AIC is lower (i.e., better model fit) for simple but RE if more biologically appropriate and does not significantly change interpretation
PSTR_bw_lm_simple <- lm(percent_change_bw ~ treatment, data = PSTR_bw)
AIC(PSTR_bw_lm, PSTR_bw_lm_simple)


# Stats output for selected model
summary(PSTR_bw_lm_simple)
anova(PSTR_bw_lm_simple) 
emmeans(PSTR_bw_lm_simple, pairwise ~ treatment, adjust = "tukey")

## Final PSTR BW model: PSTR_bw_lm_simple <- lm(percent_change_bw ~ treatment, data = PSTR_bw)

```


# Survivorship

## **NOT SURE HOW TO FIX WITH RE-IDENTIFIED SSID SAMPLES -- don't we need to know who specifically they are?** ##

## Load data and organize

```{r}
#| eval: false
#Load and format code
sur <- read.csv("mort_RT.csv")
sur<-subset(sur, mortality!='missing') #remove 2 frags that went missing (glue detached)

sur$treatment <- ifelse(sur$collection_site == "False Caye" & sur$transplant_site == "False Caye", "NS Native",
         ifelse(sur$collection_site == "False Caye" & sur$transplant_site == "Silk Cayes", "Transplant to OS",
                ifelse(sur$collection_site == "Silk Cayes" & sur$transplant_site == "Silk Cayes", "OS Native",
                       ifelse(sur$collection_site == "Silk Cayes" & sur$transplant_site == "False Caye", "Transplant to NS", NA))))

#Prepare data for graphing
#Raw data
#P. strigosa
sur_pstr <- subset(sur, species == "PSTR")

pstr.r <- Surv(sur_pstr$tod, sur_pstr$mort)
pstr.out <- survfit(pstr.r ~ treatment, data = sur_pstr)
pstr.PH <- coxme(pstr.r ~ treatment + (1 | colony_ID), sur_pstr)

#S. siderea
sur_ssid <- subset(sur, species == "SSID")

ssid.r <- Surv(sur_ssid$tod, sur_ssid$mort)
ssid.out <- survfit(ssid.r ~ treatment, data = sur_ssid)
ssid.PH <- coxme(ssid.r ~ treatment + (1 | colony_ID), sur_ssid)
summary(ssid.PH)

#Graph plot of mortality by species, transplant treatment, and time point  
treat = c("NS Native", "OS Native", "Transplant to NS", "Transplant to OS") # for treatment order
color = c("#E18A2A", "#384B6B", "#384B6B", "#E18A2A") # for colours 
size=c(12,9,4,4) #run for SSID plot
size=c(16,12,6,3) #run for PSTR plot

#P. strigosa
plot(pstr.out, col=color, xaxt="n", lwd=3, lty = c(1,1,2,2), xlab="Months after start of experiment", ylab="Fraction Surviving") 
axis(side=1, at=c(0,3,9,17, 60)) # add x axis label for time points
legend("bottomleft", col=color, lty=c(1,1,2,2), legend=treat, lwd=3, bty="n") # add legend
legend("bottomright", legend=c("p < 0.01"), bty = "n")

#S. siderea
plot(ssid.out, col=color, xaxt="n", lwd=3, lty = c(1,1,2,2), xlab="Months after start of experiment", ylab="Fraction Surviving") 
axis(side=1, at=c(0,3,9,17, 60)) # add x axis label for time points
legend("bottomleft", col=color, lty=c(1,1,2,2), legend=treat, lwd=3, bty="n") # add legend
legend("bottomright", legend=c("p < 0.01"), bty = "n")

```

## Figure creation 

### 2/24/26 Updated Mortality figure for BEM poster 
Mortality FIGURE modified for BEM Post Presentation 

Modifications: 
- make the outline of each plot match the color for each species 
- also add species name above each graph (title) instead of A and B 
- try to jitter the lines

```{r}
#Load packages
library(tidyverse)
library(ggplot2)
library(tidyr)

library(survival)
library(coxme) 
library(car)
library(tidyr)
library(dplyr)
library(survminer) 

#Import code
sur <- read.csv("mort_RT.csv")
#View(sur)

sur<-subset(sur, mortality!='missing') #remove 2 frags that went missing (glue detached)

sur$tod[is.na(sur$tod)] <- 60 # needs to be a value for the analysis (the '1' or '0' in the mort column lets the function know if there was a mortality event or not)
sur <- subset(sur, months == "60") # only want each fragment only once in the dataframe so selected the last TP

#fix the names of the transplants before we create the graphs!
sur$treatment <-
  ifelse(sur$collection_site == "False Caye" & sur$transplant_site == "False Caye", "NS Native",
      ifelse(sur$collection_site == "False Caye" & sur$transplant_site == "Silk Cayes", "Transplant to OS",
      ifelse(sur$collection_site == "Silk Cayes" & sur$transplant_site == "Silk Cayes", "OS Native",
      ifelse(sur$collection_site == "Silk Cayes" & sur$transplant_site == "False Caye", "Transplant to NS", NA))))
#View(sur)

# subset by species to perform analyses by species
sur_pstr <- subset(sur, species == "PSTR")
sur_ssid <- subset(sur, species == "SSID")

#####.  S. siderea survival plot.  ####

# Setting up data for graphing 
ssid.r <- Surv(sur_ssid$tod, sur_ssid$mort) # creates a survival object for use in survfit() function below
#View(ssid.r)
ssid.out <- survfit(ssid.r ~ treatment, data = sur_ssid) # Computes an estimate of a survival curve based on transplant location using Kaplan-Meier; this will also be used for plotting
ssid.out # view output to see mortality event and sample size summary stats

ssid.PH <- coxme(ssid.r ~ treatment + (1 | colony_ID), sur_ssid) # calculate Cox proportional hazards ratios with random effects of colony
summary(ssid.PH) # view summary of model output
anova(ssid.PH, test='Chisq') # test significance with Chi sqs for effects of transplant on survival

######### SSID GRAPH for 2/25/26 BEM MEETING ###########

# Set standards for graphing!
treat = c("NS Native", "OS Native", "Transplant to NS", "Transplant to OS") # for treatment order
color = c("#D81B60", "#004D40", "#1E88E5", "#FFC107") # for colours 
size=c(12,9,4,4) #run for SSID plot


# Graph 
quartz()
plot(ssid.out, 
     col=color, 
     xaxt="n", 
     lwd=size, 
     cex.axis = 1.4,
     xlab="Months after transplant", 
     ylab="Fraction Surviving", 
     main="SSID survival by transplant site",
     cex.lab = 1.5,    # Increase axis labels font size
     )   # Increase axis tick mark labels font size 
axis(side=1, at=c(0,3,9,17, 60), cex.axis = 1.4) # add x axis label for time points
box(col = "#E18A2A", lwd = 10) # add color to outline the box
legend("bottomleft", col=color, lty=1, legend=treat, lwd=size, bty="n", cex = 1.2) # add legend
legend("bottom", legend=c("p < 0.01"), bty = "n", cex = 2)

####. P. strigosa survival plot. ####

# Setting up data 
pstr.r <- Surv(sur_pstr$tod, sur_pstr$mort) # creates a survival object for use in survfit() function below
#View(pstr.r)
pstr.out <- survfit(pstr.r ~ treatment, data = sur_pstr) # Computes an estimate of a survival curve based on transplant location using Kaplan-Meier; this will also be used for plotting
pstr.out # view output to see mortality event and sample size summary stats

pstr.PH <- coxme(pstr.r ~ treatment + (1 | colony_ID), sur_pstr) # calculate Cox proportional hazards ratios with random effects of colony
summary(pstr.PH) # view summary of model output
anova(pstr.PH, test='Chisq') # test significance with Chi sqs for effects of transplant on survival

######### PSTR GRAPH for 2/25/26 BEM MEETING ###########

# Set standards for graphing!
treat = c("NS Native", "OS Native", "Transplant to NS", "Transplant to OS") # for treatment order
color = c("#D81B60", "#004D40", "#1E88E5", "#FFC107") # for colours 
size=c(17,12,6,3) #run for PSTR plot

# Graph 
quartz()
plot(pstr.out, 
     col=color, 
     xaxt="n", 
     lwd=size, 
     cex.axis = 1.4,
     xlab="Months after transplant", 
     ylab="Fraction Surviving", 
     main="PSTR survival by transplant site",
     cex.lab = 1.5,    # Increase axis labels font size
)   # Increase axis tick mark labels font size 
axis(side=1, at=c(0,3,9,17, 60), cex.axis = 1.4) # add x axis label for time points
box(col = "#384B6B", lwd = 10) # add color to outline the box
legend("bottomleft", col=color, lty=1, legend=treat, lwd=size, bty="n", cex = 1.2) # add legend
legend("bottom", legend=c("p < 0.01"), bty = "n", cex = 2)

```

```{r CBB surv figures}

# Save as PDF
#pdf("Final_figures/survival_Fig_Sep_17_26.pdf", width = 10, height = 4.5)
png("Final_figures/survival_Fig_Sep_17_26.png", width = 10, height = 4.5, unit = "in", res = 300)

par(mfrow = c(1, 2))  # 1 row, 2 columns

# --- Panel A: S. siderea ---
plot(ssid.out, col = color, xaxt = "n", lwd = 3, lty = c(1, 1, 2, 2),
     xlab = "Months after start of experiment",
     ylab = "Fraction Surviving",
     main = "A)  S. siderea survival")
axis(side = 1, at = c(0, 3, 9, 17, 60))
legend("bottomleft", col = color, lty = c(1, 1, 2, 2), legend = treat, lwd = 3, bty = "n")
legend("bottomright", legend = c("p < 0.01"), bty = "n")

# --- Panel B: P. strigosa ---
plot(pstr.out, col = color, xaxt = "n", lwd = 3, lty = c(1, 1, 2, 2),
     xlab = "Months after start of experiment",
     ylab = "Fraction Surviving",
     main = "B)  P. strigosa survival")
axis(side = 1, at = c(0, 3, 9, 17, 60))
legend("bottomleft", col = color, lty = c(1, 1, 2, 2), legend = treat, lwd = 3, bty = "n")
legend("bottomright", legend = c("p = 0.94"), bty = "n")

dev.off()  # close and save the file

```


## Statistical analysis

```{r}
#| eval: false
#Load data 

#Test for normality

#S. siderea by treatment
shapiro.test(subset(sur_ssid, treatment == "Transplant to OS")$mort)
shapiro.test(subset(sur_ssid, treatment == "Transplant to NS")$mort)
shapiro.test(subset(sur_ssid, treatment == "OS Native")$mort)
shapiro.test(subset(sur_ssid, treatment == "NS Native")$mort)

#P. strigosa by treatment 
shapiro.test(subset(sur_pstr, treatment == "Transplant to OS")$mort)
shapiro.test(subset(sur_pstr, treatment == "Transplant to NS")$mort)
shapiro.test(subset(sur_pstr, treatment == "OS Native")$mort)
shapiro.test(subset(sur_pstr, treatment == "NS Native")$mort)

#Test for equal variance
leveneTest(mort ~ treatment, sur_ssid)
leveneTest(mort ~ treatment, sur_pstr)

#Run ANOVAs
anova(pstr.PH, test='Chisq')

anova(ssid.PH, test='Chisq')

#Summary statistics 
mortality_pstr <-sur_pstr %>%
  group_by(treatment, mort) %>%
  summarise(
    count = n())

mortality_ssid <-sur_ssid %>%
  group_by(treatment, mort) %>%
  summarise(
    count = n())
```


# Proteins

## Load data and organize

```{r}
#| eval: false
#Load and format code
compiled_protein_assay_data <- read_csv("Raw_data/Lab Data/Proteins/compiled_protein_assay_data.csv", 
                                        col_types = cols(plate_number = col_character(), 
                                                         column = col_character(), 
                                                         replicate = col_character(), 
                                                         absorbance_595 = col_number(),
                                                         concentration_ug_ml = col_number()))

#This document is using values received from a spectrometer to calculate the concentration of proteins from each coral sample. First a protein standard curve is created to calculate the protein concentration of the coral samples. 
protein_standards<-filter(compiled_protein_assay_data, protein_source=='standard')

bsa_standard_curve <- ggplot(data=protein_standards, aes(x=absorbance_595, y=concentration_ug_ml, color=plate_number)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  xlab("Absorbance at 595") + 
  ylab("Concentration (ug/ml)") + 
  theme_classic()

#Also some coral samples required a rerun in the spectrometer. Therefore some sample replicates (with abosorbance values over those of the maximum concentration standards) were removed and added from their "redo" plate results. 

plate_3<-filter(protein_standards, plate_number==3)
plate_4<-filter(protein_standards, plate_number==4)
#Remove the weird values from each plate
  #weird values plate_3: BSA_100 rep 1, BSA_500 rep 3, BSA_750 rep 3, BSA_1000 rep 1, BSA_2000 rep 1
  #weird values plate_4: BSA_500 rep 1
plate_3<-plate_3[-c(7,15,18,19,22),]
plate_4<-plate_4[-c(13),]

#Calculate summary values for each plate's standard curve
sum_3<-summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_3))
R2_3 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_3))$r.squared
m_3 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_3))$coefficients[2,1]
b_3 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_3))$coefficients[1,1]
R2_3 #0.93166
m_3 #2247.70
b_3 #-834.28

sum_4<-summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_4))
R2_4 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_4))$r.squared
m_4 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_4))$coefficients[2,1]
b_4 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=plate_4))$coefficients[1,1]
R2_4 #0.9380535
m_4 #1976.897
b_4 #-795.9618

#Now we can have R calculate the protein concentration for each coral sample! Add these standard linear equation values into new columns in a dataframe of ONLY samples to calculate protein concentrations. 
protein_samples<-filter(compiled_protein_assay_data, protein_source=='sample')

protein_samples$m<- ifelse(protein_samples$plate_number==3, m_3, m_4)
protein_samples$b<- ifelse(protein_samples$plate_number==3, b_3, b_4)
protein_samples$calc_conc_ug_ml<- protein_samples$m * protein_samples$absorbance_595 + protein_samples$b

    # PROTEIN SAMPLE REDOS -- 5 of our samples either had 1 negative calculated protein concentration value from the above analysis or had all 3 replicates with negative values.  
#After uploading, the same code was used above to check concnetration values for redos
protein_redos_compiled_data <- read_csv("Raw_data/Lab Data/Proteins/protein_redos_compiled_data.csv", 
                                        col_types = cols(plate_number = col_character(), 
                                                         column = col_character(), protein_source = col_character(), 
                                                         replicate = col_character(), absorbance_595 = col_number(), 
                                                         concentration_ug_ml = col_number()))
protein_redo_standards<-filter(protein_redos_compiled_data, protein_source=="standard")

redo_bsa_standard_curve <- ggplot(data=protein_redo_standards, aes(x=absorbance_595, y=concentration_ug_ml, color=plate_number)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  xlab("Absorbance at 595") + 
  ylab("Concentration (ug/ml)") + 
  theme_classic()

#Remove the weird values from redo standards 
    #BSA_750 rep 3, BSA_2000 all reps
protein_redo_standards<-protein_redo_standards[-c(21,25,26,27),]

sum_5<-summary(lm(concentration_ug_ml ~ absorbance_595, data=protein_redo_standards))
R2_5 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=protein_redo_standards))$r.squared
m_5 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=protein_redo_standards))$coefficients[2,1]
b_5 <- summary(lm(concentration_ug_ml ~ absorbance_595, data=protein_redo_standards))$coefficients[1,1]
R2_5 #0.931
m_5 #953.964
b_5 #-374.905

protein_redo_samples<-filter(protein_redos_compiled_data, protein_source=='sample')
protein_redo_samples$m<- m_5
protein_redo_samples$b<- b_5
protein_redo_samples$calc_conc_ug_ml<- protein_redo_samples$m * protein_redo_samples$absorbance_595 + protein_redo_samples$b

#Combine the protein samples and the redo protein samples 
df_final_protein<-protein_samples[-c(49,50,51,58,59,60,61,62,63,70,71,72,82,83,84),]
df_final_protein2<-rbind(df_final_protein, protein_redo_samples)

#Need to fix the name of PSTRS14 -- in there as 514!
df_final_protein2$sample_ID<-ifelse(df_final_protein2$sample_ID == 'PSTR514_1', 'PSTRS14_1', 
                                    as.character(df_final_protein2$sample_ID))
#View(df_final_protein2)

#Prepare data for graphing

  #Accounting for volume of tissue slurry in calculated protein concentrations 
energy_starting_volume<- read_csv("Raw_data/Lab Data/Proteins/energyreserve_starting_volume.csv", 
                                  col_types = cols(total_tissue_slurry_volume_ml = col_number()))
final_protein2<-merge(df_final_protein2, energy_starting_volume, by="sample_ID")

protein_abs_means<- final_protein2 %>%
  group_by(sample_ID) %>%
  summarise(mean_pro=mean(calc_conc_ug_ml), 
            sd=sd(calc_conc_ug_ml), 
            n=n(), 
            se=sd/sqrt(n), 
            meanvol=mean(total_tissue_slurry_volume_ml), 
            m=mean(m), 
            b=mean(b))


# protein_abs_means<- final_protein2 %>%
#   group_by(sample_ID) %>%
#   summarise(meanabsor=mean(absorbance_595), 
#             sd=sd(absorbance_595), 
#             n=n(), 
#             se=sd/sqrt(n), 
#             meanvol=mean(total_tissue_slurry_volume_ml), 
#             m=mean(m), 
#             b=mean(b))

#protein_abs_means$calc_conc_ug_ml2<- protein_abs_means$m * protein_abs_means$meanabsor + protein_abs_means$b
protein_abs_means$final_protein_ug<- protein_abs_means$mean_pro * protein_abs_means$meanvol 

  #converting from ug to mg for direct comparsion to Baumann et al. 2021
protein_abs_means$final_protein_mg<- protein_abs_means$final_protein_ug / 1000


#Back Calculations to tissue biomass (mg) and surface area (cm^2)
  # only the calculations to biomass were used in the final paper 
  # (therefore, surface area associated code is commented out below)


  # (1) Back-calculating protein concentrations to biomass (mg) 
BM_measurements <- read_csv("Raw_data/Lab Data/backcalcs/BM_measurements.csv", 
                            col_types = cols(ash_free_dry_weight_g = col_number(), 
                                             dry_weight_volume_ml = col_number(), 
                                             total_volume_ml = col_number(), total_dry_weight_g = col_number()))

  #convert dry weight from g to mg for direct comparsion to Baumann et al. 2021
BM_measurements$total_dry_weight_mg<- BM_measurements$total_dry_weight_g * 1000

  #Merge data frames together 
protein_bm_analysis<-merge(protein_abs_means, BM_measurements, by="sample_ID")

  #Calculate final concentration of proteins in mg/mg for direct comparsion to Baumann et al. 2021
protein_bm_analysis$protein_mg_mg<- protein_bm_analysis$final_protein_mg / protein_bm_analysis$total_dry_weight_mg

#Raw data
samples_treatments <- read_csv("Raw_data/Lab Data/backcalcs/samples_treatments.csv")
protein_mg_mg_treatments<-merge(protein_bm_analysis, samples_treatments, by="sample_ID")

  #fixing transplant treatment names
protein_mg_mg_treatments$treatment<-ifelse(protein_mg_mg_treatments$treatment == 'Native NS', 'NS Native', as.character(protein_mg_mg_treatments$treatment))
protein_mg_mg_treatments$treatment<-ifelse(protein_mg_mg_treatments$treatment == 'Native OS', 'OS Native', as.character(protein_mg_mg_treatments$treatment))

  #save cleaned protein concentration (to biomass) data for records
write_csv(protein_mg_mg_treatments, file="Cleaned_data/protein_data.csv")
#write_xlsx(protein_mg_mg_treatments, path="~/Desktop/Desktop - Callie’s MacBook Pro/Undergrad Research /Honors Thesis/Publication/Cleaned_data/protein_data.xlsx")


  # (2) Back-calculating protein concentrations to surface area (cm^2)
#Import data set
# sa_measurements <- read_csv("Lab Data/backcalcs/sa_measurements.csv", 
#     col_types = cols(standard_area_cm2 = col_number(), 
#         avg_standard_area_pixels = col_number(), 
#         avg_sa_fragment_pixels = col_number(),
#         sample_area_cm2 = col_number()))
# #View(sa_measurements)
# 
# #combine the datasets!
# protein_sa_analysis<-merge(protein_abs_means, sa_measurements, by="sample_ID")
# #View(protein_sa_analysis)
# 
# #now calculate mg/cm^2 -- type out the calculation necessary for this!
# protein_sa_analysis$protein_mg_cm2<- protein_sa_analysis$final_protein_mg / protein_sa_analysis$sample_area_cm2
# #View(protein_sa_analysis)
# 
#   #save carb concentration (mg/cm^2) data for records 
# write_csv(protein_sa_analysis, file="all_corals_protein_mgcm2_data.csv")
# write_xlsx(protein_sa_analysis, path="all_corals_protein_mgcm2_data.xlsx")

```

## Figure creation

```{r}
  #removing corals missing treatment ID for graphing 
protein_mg_mg_treatments_noNA<-filter(protein_mg_mg_treatments, treatment!="na")

  #adding differentiation of environment assignments for panels in graph  
protein_mg_mg_treatments_noNA<- protein_mg_mg_treatments_noNA %>% 
  mutate(transplant_envir = ifelse(grepl("NS", treatment), "Nearshore", "Offshore"))
protein_mg_mg_treatments_noNA<- protein_mg_mg_treatments_noNA %>% 
  mutate(home_envir = ifelse(treatment == "NS Native" | treatment == "Transplant to OS", 
                             "Nearshore coral", "Offshore coral"))

#Mean data
Protein_BM_mean_plot_data <- protein_mg_mg_treatments_noNA %>%
  group_by(species, treatment, transplant_envir, home_envir) %>%
  summarise(mean=mean(protein_mg_mg), sd=sd(protein_mg_mg), n=n(), se=sd/sqrt(n))


## CBB edit: what you were trying above:
Protein_BM_mean_plot_data$species <- factor(Protein_BM_mean_plot_data$species, levels = c("SSID","PSTR"))
protein_mg_mg_treatments_noNA$species <- factor(protein_mg_mg_treatments_noNA$species, levels = c("SSID","PSTR"))


#Graph plot of protein by species and transplant treatment 

Protein_Fig <- ggplot(Protein_BM_mean_plot_data %>% drop_na(), 
       aes(x=treatment, y=mean, color=species))+
  geom_point(data=protein_mg_mg_treatments_noNA, 
             aes(x=treatment, y=protein_mg_mg, color=species), 
             position=pd, size=2.5, alpha=0.3)+
  geom_point(size=3, position=pd)+
  geom_line(aes(group=species), position=pd)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.2, size=1, position=pd)+
  facet_wrap(~home_envir, scales='free_x')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), # make axes title text larger
        axis.text = element_text(size = 12), # make axes text larger  
        legend.title = element_text(size = 13), # make legend title larger 
        legend.text = element_text(size = 13, face = "italic"), # make legend text larger
        strip.text = element_text(size = 13)) +  # make text in facet title larger
  guides(color = guide_legend(title = "Species")) + 
  xlab("Treatment") +
  ylab("Mean [protein] (mg/mg)")+
  scale_color_manual(labels=c('P. strigosa', 'S. siderea'), values = c("#384B6B", "#E18A2A"))


Protein_Fig2 <- ggplot(data = Protein_BM_mean_plot_data %>% drop_na(), aes(x = transplant_envir, y=mean, color = home_envir, group = home_envir)) +
  geom_errorbar(aes(x=transplant_envir, ymin=mean-se, ymax=mean+se), width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = protein_mg_mg_treatments_noNA, aes(x=transplant_envir, y=protein_mg_mg), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~species, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c("NS", "OS")) +
  scale_x_discrete(labels = c("NS", "OS")) +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    #legend.title = element_blank(), # remove legend title 
    strip.text = element_text(size = 14, face = "italic")) +  # make text in facet title larger & italic
  guides(color = guide_legend(title = "Native Environment")) + 
  xlab("Transplant Environment") +
  ylab("Mean [protein] (mg/mg)")


# save image as a .jpeg
ggsave("Protein_Fig_Sep_17_26.jpg", plot=Protein_Fig2, path="Final_figures", width = 10, height = 5, units = "in", dpi = 600)

#Save graphed raw data (no NAs in transplant treatment) for statistical analysis 
#write_csv(protein_mg_mg_treatments_noNA, file="Cleaned_data/protein_noNA_data.csv")

```

## Statistical analysis for proteins backcalculated to biomass (mg/mg)

```{r}
#| eval: false
#Load data 
protein_bm_data <- read_csv("Cleaned_data/protein_noNA_data.csv",
                            col_types = cols(protein_mg_mg = col_number()))

#Test for normality

#S. siderea by treatment
SSID_protein_bm<-subset(protein_bm_data, species == "SSID")

shapiro.test(subset(SSID_protein_bm, treatment == "Transplant to NS")$protein_mg_mg)
  # p-value > 0.05
#no SSID transplant to OS corals
shapiro.test(subset(SSID_protein_bm, treatment == "OS Native")$protein_mg_mg)
  # p-value > 0.05
shapiro.test(subset(SSID_protein_bm, treatment == "NS Native")$protein_mg_mg)
  # p-value > 0.05

#P. strigosa by treatment  
PSTR_protein_bm<-subset(protein_bm_data, species == "PSTR")

#one PSTR transplant to NS coral
shapiro.test(subset(PSTR_protein_bm, treatment == "Transplant to OS")$protein_mg_mg)
  # p-value < 0.05
shapiro.test(subset(PSTR_protein_bm, treatment == "OS Native")$protein_mg_mg)
  # p-value > 0.05
shapiro.test(subset(PSTR_protein_bm, treatment == "NS Native")$protein_mg_mg)
  # p-value > 0.05

#Test for equal variance
leveneTest(protein_mg_mg ~ treatment, SSID_protein_bm)
  # p-value > 0.05
leveneTest(protein_mg_mg ~ treatment, PSTR_protein_bm)
  # p-value > 0.05

#Run ANOVAs
#======comparing species=======#
aovp1<-aov(protein_mg_mg~species, data=protein_bm_data)
summary(aovp1)

#======comparing treatments for each species=======#
  #S. siderea
aovps<-aov(protein_mg_mg~treatment, data=SSID_protein_bm)
summary(aovps)
  #P. strigosa
aovpp<-aov(protein_mg_mg~treatment, data=PSTR_protein_bm)
summary(aovpp)

#Summary statistics 
stats_proteins <-protein_bm_data %>%
  group_by(species, treatment) %>%
  summarise(
          count = n(),
          mean_protein = mean(protein_mg_mg))
stats_proteins
```


# Carbohydrates

## Load data and organize

```{r}
#| eval: false
#Load and format code
Carb_data <- read_csv("Raw_data/Lab Data/Carbohydrates/Carbohydrate Assays - compiled_carb_assay_data.csv", 
                      col_types = cols(plate_number = col_character(), 
                                       row = col_character(), 
                                       column = col_character(), 
                                       replicate = col_character(), 
                                       absorbance_485 = col_number(), 
                                       concentration_mg_ml = col_number()))

#This document is using values received from a spectrometer to calculate the concentration of carbohydrates from each coral sample. First a carb standard curve is created to calculate the carb concentration of the coral samples. 
carb_standards<-filter(Carb_data, carb_source=='standard')

glucose_standard_curve <- ggplot(data=carb_standards, aes(x=absorbance_485, y=concentration_mg_ml, color=plate_number)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  xlab("Absorbance at 485") + 
  ylab("Concentration (mg/ml)") + 
  theme_classic()
glucose_standard_curve

#Calculate summary values for each plate's standard curve
cplate_4<-filter(carb_standards, plate_number==4)
cplate_5<-filter(carb_standards, plate_number==5)

sum_cplate4<-summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_4))
R2_cplate4 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_4))$r.squared
m_cplate4 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_4))$coefficients[2,1]
b_cplate4 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_4))$coefficients[1,1]
R2_cplate4 #0.9979
m_cplate4 #0.8672
b_cplate4 #-0.211298

sum_cplate5<-summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_5))
R2_cplate5 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_5))$r.squared
m_cplate5 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_5))$coefficients[2,1]
b_cplate5 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=cplate_5))$coefficients[1,1]
R2_cplate5 #0.99679
m_cplate5 #0.809757
b_cplate5 #-0.191055

#Now we can have R calculate the carbohydrate concentration for each coral sample! Add these standard linear equation values into new columns in a dataframe of ONLY samples to calculate carbohydrate concentrations. 
carb_samples<-filter(Carb_data, carb_source=='sample')
carb_samples$dilution<-(0.1)
carb_samples$m<- ifelse(carb_samples$plate_number==4, m_cplate4, m_cplate5)
carb_samples$b<- ifelse(carb_samples$plate_number==4, b_cplate4, b_cplate5)

#Also some coral samples required a rerun in the spectrometer. Therefore some sample replicates (with abosorbance values over those of the maximum concentration standards) were removed and added from their "redo" plate results. 
redo_carb_data <- read_csv("Raw_data/Lab Data/Carbohydrates/Carbohydrate Assays - compiled_carb_redo_data.csv", col_types = cols(plate_number = col_character(), column = col_character(), replicate = col_character(), absorbance_485 = col_number(), concentration_mg_ml = col_number()))

redo_carb_standards<-filter(redo_carb_data, carb_source=='standard')

glucose_redo_standard_curve <- ggplot(data=redo_carb_standards, aes(x=absorbance_485, y=concentration_mg_ml)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  xlab("Absorbance at 485") + 
  ylab("Concentration (mg/ml)") + 
  theme_classic()
glucose_redo_standard_curve

sum_cplate6<-summary(lm(concentration_mg_ml ~ absorbance_485, data=redo_carb_standards))
R2_cplate6 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=redo_carb_standards))$r.squared
m_cplate6 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=redo_carb_standards))$coefficients[2,1]
b_cplate6 <- summary(lm(concentration_mg_ml ~ absorbance_485, data=redo_carb_standards))$coefficients[1,1]
R2_cplate6 #0.99211
m_cplate6 #0.920258
b_cplate6 #-0.24104

redo_carb_samples<-filter(redo_carb_data, carb_source=='sample')
redo_carb_samples$m<- R2_cplate6
redo_carb_samples$b<- b_cplate6

#Combine the carb samples and the redo carb samples 

#First remove redo samples (SSIDF20_3 and SSID42_1) out of the carb_samples data frame
carb_samples2<-carb_samples[-c(7,8,9,28,29,30),-c(1)]
#Pick which dilution redo using for each of these from redo 
redo_carb_samples2<-redo_carb_samples[-c(1,2,3,10,11,12),]
redo_carb_samples3<-redo_carb_samples2[,-c(5)]
redo_carb_samples3$dilution<- ifelse(redo_carb_samples3$sample_ID=="SSIDF20_3", 0.083, 0.067)
#View(redo_carb_samples3)

#combine data sets together
df_final_carb<-rbind(carb_samples2, redo_carb_samples3)

#Prepare data for graphing
  #Accounting for volume of tissue slurry in calculated carbohydrate concentrations
energy_starting_volume2 <- read_csv("Raw_data/Lab Data/Proteins/energyreserve_starting_volume.csv", 
                                          col_types = cols(total_tissue_slurry_volume_ml = col_number()))
final_carb<-merge(df_final_carb, energy_starting_volume2, by="sample_ID")
  
  #mean value for concentration (mg/mL) for each sample 
carb_abs_means<- final_carb %>%
  group_by(sample_ID) %>%
  summarise(meanabsor=mean(absorbance_485), sd=sd(absorbance_485), n=n(), se=sd/sqrt(n), meanvol=mean(total_tissue_slurry_volume_ml), m=mean(m), b=mean(b), dilution=mean(dilution))

carb_abs_means$calc_conc_mg_ml<- carb_abs_means$m * carb_abs_means$meanabsor + carb_abs_means$b

  # account for starting volume to calculate concentration in mg
carb_abs_means$final_carb_mg<- carb_abs_means$calc_conc_mg_ml * (carb_abs_means$dilution * carb_abs_means$meanvol) 

  # (1) Back-calculating carbohydrate concentrations to biomass (g) 
BM_measurements <- read_csv("Raw_data/Lab Data/backcalcs/BM_measurements.csv", 
                            col_types = cols(ash_free_dry_weight_g = col_number(), 
                                             dry_weight_volume_ml = col_number(), 
                                             total_volume_ml = col_number(), total_dry_weight_g = col_number()))
carb_bm_analysis<-merge(carb_abs_means, BM_measurements, by="sample_ID")

carb_bm_analysis$carb_mg_mg<- (carb_bm_analysis$final_carb_mg / (carb_bm_analysis$total_dry_weight_g*1000))

#Raw data
samples_treatments <- read_csv("Raw_data/Lab Data/backcalcs/samples_treatments.csv")
carb_mgmg_treatments<-merge(carb_bm_analysis, samples_treatments, by="sample_ID")
  
  #fixing treatment names
carb_mgmg_treatments$treatment<-ifelse(carb_mgmg_treatments$treatment == 'Native NS', 'NS Native', 
                                as.character(carb_mgmg_treatments$treatment))
carb_mgmg_treatments$treatment<-ifelse(carb_mgmg_treatments$treatment == 'Native OS', 'OS Native', 
                                 as.character(carb_mgmg_treatments$treatment))


  #save carb concentration (mg/mg) data for records
write_csv(carb_mgmg_treatments, file="Cleaned_data/carb_data.csv")
#write_xlsx(carb_mgmg_treatments, path="carb_data.xlsx")



  # (2) Back-calculating carbohydrate concentrations to surface area (mg/cm^2) 
#Import data set
# sa_measurements <- read_csv("Lab Data/backcalcs/sa_measurements.csv", 
#     col_types = cols(standard_area_cm2 = col_number(), 
#         avg_standard_area_pixels = col_number(), 
#         avg_sa_fragment_pixels = col_number(),
#         sample_area_cm2 = col_number()))
# #View(sa_measurements)
# 
# #combine the datasets!
# carb_sa_analysis<-merge(carb_abs_means, sa_measurements, by="sample_ID")
# #View(carb_sa_analysis)
# 
# #now calculate mg/cm^2 -- type out the calculation necessary for this!
# carb_sa_analysis$carb_mg_cm2<- carb_sa_analysis$final_carb_mg / carb_sa_analysis$sample_area_cm2
# #View(carb_sa_analysis)
# 
# 
#   #save carb concentration (mg/cm^2) data for records 
# write_csv(carb_sa_analysis, file="all_corals_carb_mgcm2_data.csv")
# write_xlsx(carb_sa_analysis, path="all_corals_carb_mgcm2_data.xlsx")

```

## Figure creation

```{r}
#Graphing back calculated carbs to biomass (mg/mg)

  #removing corals missing treatment ID
carb_mgmg_treatments_noNA<-filter(carb_mgmg_treatments, treatment!="na")

  #adding differentiation of environment assignments for panels in graph
carb_mgmg_treatments_noNA<- carb_mgmg_treatments_noNA %>% 
  mutate(transplant_envir = ifelse(grepl("NS", treatment), "Nearshore", "Offshore"))
carb_mgmg_treatments_noNA<- carb_mgmg_treatments_noNA %>% 
  mutate(home_envir = ifelse(treatment == "NS Native" | treatment == "Transplant to OS", 
                             "Nearshore coral", "Offshore coral"))

#Mean data
Carbs_BM_mean_plot_data <- carb_mgmg_treatments_noNA %>%
  group_by(species, treatment, transplant_envir, home_envir) %>%
  summarise(mean=mean(carb_mg_mg), sd=sd(carb_mg_mg), n=n(), se=sd/sqrt(n))
#View(Carbs_BM_mean_plot_data)


## CBB edit: what you were trying above:
Carbs_BM_mean_plot_data$species <- factor(Carbs_BM_mean_plot_data$species, levels = c("SSID","PSTR"))
carb_mgmg_treatments_noNA$species <- factor(carb_mgmg_treatments_noNA$species, levels = c("SSID","PSTR"))

#Graph plot of carbohydrates by species and transplant treatment 

Carbs_Fig <- ggplot(Carbs_BM_mean_plot_data, 
                        aes(x=treatment, y=mean, color=species))+
  geom_point(data=carb_mgmg_treatments_noNA, 
             aes(x=treatment, y=carb_mg_mg, color=species), 
             position=pd, size=2.5, alpha=0.3)+
  geom_point(size=3, position=pd)+
  geom_line(aes(group=species), position=pd)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.2, size=1, position=pd)+
  facet_wrap(~home_envir, scales='free_x') +
  theme_bw() + 
  theme(axis.title = element_text(size = 14), # make axes title text larger
        axis.text = element_text(size = 12), # make axes text larger  
        legend.title = element_text(size = 13), # make legend title larger 
        legend.text = element_text(size = 13, face = "italic"), # make legend text larger
        strip.text = element_text(size = 13)) +  # make text in facet title larger
  guides(color = guide_legend(title = "Species")) + 
  xlab("Treatment") +
  ylab("Mean [carbohydrate] (mg/mg)")+
  scale_color_manual(labels=c('P. strigosa', 'S. siderea'), values = c("#384B6B", "#E18A2A"))




Carbs_Fig2 <- ggplot(data = Carbs_BM_mean_plot_data %>% drop_na(), aes(x = transplant_envir, y=mean, color = home_envir, group = home_envir)) +
  geom_errorbar(aes(x=transplant_envir, ymin=mean-se, ymax=mean+se), width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = carb_mgmg_treatments_noNA, aes(x=transplant_envir, y=carb_mg_mg), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~species, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c("NS", "OS")) +
  scale_x_discrete(labels = c("NS", "OS")) +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    #legend.title = element_blank(), # remove legend title 
    strip.text = element_text(size = 14, face = "italic")) +  # make text in facet title larger & italic
  guides(color = guide_legend(title = "Native Environment")) + 
  xlab("Transplant Environment") +
  ylab("Mean [carbohydrate] (mg/mg)")



# save image as a .jpeg
ggsave("Carb_Fig_Sep_17_26.jpg", plot=Carbs_Fig2, path="Final_figures", width = 10, height = 5, units = "in", dpi = 600)

#Save graphed raw data (w/no NA for treatment) for statistical analysis 
#write_csv(carb_mgmg_treatments_noNA, file="Cleaned_data/carbs_noNA_data.csv")
```


## Statistical analysis

```{r}
#| eval: false
#Load data 
carbs_bm_data <- read_csv("Cleaned_data/carbs_noNA_data.csv",
                            col_types = cols(carb_mg_mg = col_number()))
#Test for normality

#S. siderea by treatment
SSID_carbs_bm<-subset(carbs_bm_data, species == "SSID")

shapiro.test(subset(SSID_carbs_bm, treatment == "Transplant to NS")$carb_mg_mg)
  # p-value > 0.05
#no SSID transplant to OS corals
shapiro.test(subset(SSID_carbs_bm, treatment == "OS Native")$carb_mg_mg)
  # p-value > 0.05
shapiro.test(subset(SSID_carbs_bm, treatment == "NS Native")$carb_mg_mg)
  # p-value > 0.05

#P. strigosa by treatment 
PSTR_carbs_bm<-subset(carbs_bm_data, species == "PSTR")

#one PSTR transplant to NS coral
shapiro.test(subset(PSTR_carbs_bm, treatment == "Transplant to OS")$carb_mg_mg)
  # p-value > 0.05
shapiro.test(subset(PSTR_carbs_bm, treatment == "OS Native")$carb_mg_mg)
  # p-value > 0.05
shapiro.test(subset(PSTR_carbs_bm, treatment == "NS Native")$carb_mg_mg)
  # p-value > 0.05

#Test for equal variance
leveneTest(carb_mg_mg ~ treatment, SSID_carbs_bm)
  # p-value > 0.05
leveneTest(carb_mg_mg ~ treatment, PSTR_carbs_bm)
  # p-value > 0.05

#Run ANOVAs + Tukey HSD
#======comparing species=======#
aovc1<-aov(carb_mg_mg~species, data=carbs_bm_data)
summary(aovc1)
TukeyHSD(aovc1)

#======comparing treatments for each species=======#
  #S. siderea
aovcs<-aov(carb_mg_mg~treatment, data=SSID_carbs_bm)
summary(aovcs)
  #P. strigosa
aovcp<-aov(carb_mg_mg~treatment, data=PSTR_carbs_bm)
summary(aovcp)
TukeyHSD(aovcp)

#Summary statistics 
stats_carbs <-carbs_bm_data %>%
  group_by(species, treatment) %>%
  summarise(
          count = n(),
          mean_carbs = mean(carb_mg_mg))
stats_carbs
```


# Lipids

## Load data and organize

```{r}
#| eval: false
#Load data set

Lipid_data <- read_csv("Raw_data/Lab Data/Lipids/compiled_lipid_data.csv", 
    col_types = cols(plate_number = col_character(), 
        column = col_character(), replicate = col_character(), 
        blank_correction_absorbance_540 = col_number(), 
        concentration_mg_ml = col_number()))
#View(Lipid_data)

#This document is using values received from a spectrometer to calculate the concentration of lipids from each coral sample. First a lipid standard is created from plotting the absorbance of known lipid concentrations (standards)
lipid_standards<-filter(Lipid_data, lipid_source=='standard')

corn_standard_curve <- ggplot(data=lipid_standards, aes(x=blank_correction_absorbance_540, y=concentration_mg_ml, color=plate_number)) + geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  xlab("Absorbance at 540") +
  ylab("Concentration (mg/ml)") +
  theme_classic() 
corn_standard_curve

#Calculate summary values for each plate's standard curve
lplate_1<-filter(lipid_standards, plate_number==1) 
lplate_2<-filter(lipid_standards, plate_number==2)

sum_lplate1<-summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_1))
R2_lplate1 <- summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_1))$r.squared 
m_lplate1 <- summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_1))$coefficients[2,1] 
b_lplate1 <- summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_1))$coefficients[1,1] 

R2_lplate1 #0.95960
m_lplate1 #1.7596
b_lplate1 #-0.019183

sum_lplate2<-summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_2))
R2_lplate2 <- summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_2))$r.squared 
m_lplate2<-summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_2))$coefficients[2,1] 
b_lplate2 <- summary(lm(concentration_mg_ml ~ blank_correction_absorbance_540, data=lplate_2))$coefficients[1,1] 

R2_lplate2 #0.932895
m_lplate2 #1.34222
b_lplate2 #0.009980

#Now we can have R calculate the lipid concentration for each coral sample! Add these standard linear equation values into new columns in a dataframe of samples to estimate lipid concentration
lipid_samples<-filter(Lipid_data, lipid_source=='sample') 
lipid_samples$dilution<-(1360/600)
lipid_samples$m<- ifelse(lipid_samples$plate_number==1, m_lplate1, m_lplate2) 
lipid_samples$b<- ifelse(lipid_samples$plate_number==1, b_lplate1, b_lplate2)

#Prepare sample data for graphing
#Accounting for volume of tissue slurry in calculated carbohydrate concentrations
energy_starting_volume <- read_csv("Raw_data/Lab Data/Proteins/energyreserve_starting_volume.csv", 
                                    col_types = cols(total_tissue_slurry_volume_ml = col_number()))
final_lipid<-merge(lipid_samples, energy_starting_volume, by="sample_ID")

lipid_abs_means<- final_lipid %>%
  group_by(sample_ID) %>%
  summarise(meanabsor=mean(blank_correction_absorbance_540), sd=sd(blank_correction_absorbance_540), n=n(), se=sd/sqrt(n), meanvol=mean(total_tissue_slurry_volume_ml), m=mean(m), b=mean(b), dilution=mean(dilution))

#calculate lipid concentrations
lipid_abs_means$calc_conc_mg_ml<- lipid_abs_means$m * lipid_abs_means$meanabsor + lipid_abs_means$b
lipid_abs_means$final_lipid_mg<- lipid_abs_means$calc_conc_mg_ml * lipid_abs_means$meanvol * lipid_abs_means$dilution  

  # (1) Back-calculating lipid concentrations to biomass (g) 
BM_measurements <- read_csv("Raw_data/Lab Data/backcalcs/BM_measurements.csv", 
                            col_types = cols(ash_free_dry_weight_g = col_number(), 
                                             dry_weight_volume_ml = col_number(), 
                                             total_volume_ml = col_number(), total_dry_weight_g = col_number()))

#combine datasets
lipid_bm_analysis<-merge(lipid_abs_means, BM_measurements, by="sample_ID")

#calculate concentration mg/mg
lipid_bm_analysis$lipid_mg_mg<- lipid_bm_analysis$final_lipid_mg / (lipid_bm_analysis$total_dry_weight_g*1000)

  #assign treatments to samples
samples_treatments <- read_csv("Raw_data/Lab Data/backcalcs/samples_treatments.csv")

lipid_mgmg_treatments<-merge(lipid_bm_analysis, samples_treatments, by="sample_ID")

  #fixing treatment names
lipid_mgmg_treatments$treatment<-ifelse(lipid_mgmg_treatments$treatment == 'Native NS', 'NS Native', 
                                       as.character(lipid_mgmg_treatments$treatment))
lipid_mgmg_treatments$treatment<-ifelse(lipid_mgmg_treatments$treatment == 'Native OS', 'OS Native', 
                                       as.character(lipid_mgmg_treatments$treatment))


#save lipid concentration (mg/mg) data for records
write_csv(lipid_mgmg_treatments, file="Cleaned_data/lipid_data.csv")
#write_xlsx(lipid_mgmg_treatments, path="~/Cleaned_data/lipid_data.xlsx")

  # (2) Back-calculating lipid concentrations to surface area (cm^2)
# sa_measurements <- read_csv("Lab Data/backcalcs/sa_measurements.csv", 
#                             col_types = cols(standard_area_cm2 = col_number(), 
#                                              avg_standard_area_pixels = col_number(), 
#                                              avg_sa_fragment_pixels = col_number(),
#                                              sample_area_cm2 = col_number()))
# #View(sa_measurements)
# 
# #combine the datasets!
# lipid_sa_analysis<-merge(lipid_abs_means, sa_measurements, by="sample_ID")
# #View(lipid_sa_analysis)
# 
# #calculate concentration mg/cm2
# lipid_sa_analysis$lipid_mg_cm2<- lipid_sa_analysis$final_lipid_mg / lipid_sa_analysis$sample_area_cm2
# #View(lipid_sa_analysis)
# 
#   #save lipid concentration (mg/cm2) data for records
# write_csv(lipid_sa_analysis, file="all_corals_lipid_mgcm2_data.csv")
# write_xlsx(lipid_sa_analysis, path="all_corals_lipid_mgcm2_data.xlsx")

```

## Figure creation 

```{r}
#Graphing back calculated lipids to biomass (mg/mg)

#removing corals missing treatment ID
lipid_mgmg_treatments_noNA<-filter(lipid_mgmg_treatments, treatment!="na")

 #adding differentiation of environment assignments for panels in graph
lipid_mgmg_treatments_noNA<- lipid_mgmg_treatments_noNA %>% 
  mutate(transplant_envir = ifelse(grepl("NS", treatment), "Nearshore", "Offshore"))
lipid_mgmg_treatments_noNA<- lipid_mgmg_treatments_noNA %>% 
  mutate(home_envir = ifelse(treatment == "NS Native" | treatment == "Transplant to OS", 
                             "Nearshore coral", "Offshore coral"))

#Mean data
Lipids_BM_mean_plot_data <- lipid_mgmg_treatments_noNA %>%
  group_by(species, treatment, transplant_envir, home_envir) %>%
  summarise(mean=mean(lipid_mg_mg), sd=sd(lipid_mg_mg), n=n(), se=sd/sqrt(n))
#View(Lipids_BM_mean_plot_data)

## CBB edit: what you were trying above:
Lipids_BM_mean_plot_data$species <- factor(Lipids_BM_mean_plot_data$species, levels = c("SSID","PSTR"))
lipid_mgmg_treatments_noNA$species <- factor(lipid_mgmg_treatments_noNA$species, levels = c("SSID","PSTR"))


#Graph plot of carbohydrates by species and transplant treatment 

Lipids_Fig <- ggplot(Lipids_BM_mean_plot_data, 
       aes(x=treatment, y=mean, color=species))+
  geom_point(data=lipid_mgmg_treatments_noNA, 
             aes(x=treatment, y=lipid_mg_mg, color=species), 
             position=pd, size=2.5, alpha=0.3)+
  geom_point(size=3, position=pd)+
  geom_line(aes(group=species), position=pd)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.2, size=1, position=pd)+
  facet_wrap(~home_envir, scales='free_x')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), # make axes title text larger
        axis.text = element_text(size = 12), # make axes text larger  
        legend.title = element_text(size = 13), # make legend title larger 
        legend.text = element_text(size = 13, face = "italic"), # make legend text larger
        strip.text = element_text(size = 13)) +  # make text in facet title larger
  guides(color = guide_legend(title = "Species")) + 
  xlab("Treatment") +
  ylab("Mean [lipid] (mg/mg)")+
  scale_color_manual(labels=c('P. strigosa', 'S. siderea'), values = c("#384B6B", "#E18A2A"))





Lipids_Fig2 <- ggplot(data = Lipids_BM_mean_plot_data %>% drop_na(), aes(x = transplant_envir, y=mean, color = home_envir, group = home_envir)) +
  geom_errorbar(aes(x=transplant_envir, ymin=mean-se, ymax=mean+se), width=0.2, linewidth=1, position=pd) +
  geom_line(position=pd) +
  geom_point(size=3, position=pd) +
  geom_point(data = lipid_mgmg_treatments_noNA, aes(x=transplant_envir, y=lipid_mg_mg), position=pjd, size=2.5, alpha=0.3) +
  facet_wrap(~species, scales='free_x', labeller = labeller(species = c("SSID" = "S. siderea", "PSTR" = "P. strigosa"))) +
  scale_color_manual(values = c("#E18A2A", "#384B6B"), labels=c("NS", "OS")) +
  scale_x_discrete(labels = c("NS", "OS")) +
  theme_bw() + 
  theme(
    axis.title = element_text(size = 15), # make axes title text larger
    axis.text = element_text(size = 12), # make axes text larger  
    #legend.title = element_blank(), # remove legend title 
    strip.text = element_text(size = 14, face = "italic")) +  # make text in facet title larger & italic
  guides(color = guide_legend(title = "Native Environment")) + 
  xlab("Transplant Environment") +
  ylab("Mean [lipid] (mg/mg)")

# save image as a .jpeg
ggsave("Lipid_Fig_Sep_17_26.jpg", plot=Lipids_Fig2, path="Final_figures", width = 10, height = 5, units = "in", dpi = 600)

#Save graphed raw data for statistical analysis 
#write_csv(lipid_mgmg_treatments_noNA, file="Cleaned_data/lipid_noNA_data.csv")
```


## Statistical analysis

```{r}
#| eval: false
#Load data 
lipids_bm_data <- read_csv("Cleaned_data/lipid_noNA_data.csv",
                            col_types = cols(lipid_mg_mg = col_number()))

#Test for normality

#S. siderea by treatment
SSID_lipids_bm<-subset(lipids_bm_data, species == "SSID")

shapiro.test(subset(SSID_lipids_bm, treatment == "Transplant to NS")$lipid_mg_mg)
  # p-value > 0.05
#no SSID transplant to OS corals
shapiro.test(subset(SSID_lipids_bm, treatment == "OS Native")$lipid_mg_mg)
  # p-value > 0.05
shapiro.test(subset(SSID_lipids_bm, treatment == "NS Native")$lipid_mg_mg)
  # p-value > 0.05

#P. strigosa by treatment 
PSTR_lipids_bm<-subset(lipids_bm_data, species == "PSTR")

#one PSTR transplant to NS coral
shapiro.test(subset(PSTR_lipids_bm, treatment == "Transplant to OS")$lipid_mg_mg)
  # p-value > 0.05
shapiro.test(subset(PSTR_lipids_bm, treatment == "OS Native")$lipid_mg_mg)
  # p-value > 0.05
shapiro.test(subset(PSTR_lipids_bm, treatment == "NS Native")$lipid_mg_mg)
  # p-value < 0.05

#Test for equal variance
leveneTest(lipid_mg_mg ~ treatment, SSID_lipids_bm)
  # p-value > 0.05
leveneTest(lipid_mg_mg ~ treatment, PSTR_lipids_bm)
  # p-value > 0.05

#Run ANOVAs + Tukey HSD
#======comparing species=======#
aovc1<-aov(lipid_mg_mg~species, data=lipids_bm_data)
summary(aovc1)


#======comparing treatments for each species=======#
  #S. siderea
aovcs<-aov(lipid_mg_mg~treatment, data=SSID_lipids_bm)
summary(aovcs)
  #P. strigosa
aovcp<-aov(lipid_mg_mg~treatment, data=PSTR_lipids_bm)
summary(aovcp)


#Summary statistics 
stats_lipids <-lipids_bm_data %>%
  group_by(species, treatment) %>%
  summarise(
          count = n(),
          mean_carbs = mean(lipid_mg_mg))
stats_lipids
```

